---
title: "US Agricultural Tariffs: A Comprehensive Analysis"
subtitle: "USITC Tariff Database · HTS Chapters 1–24 · TidyTuesday 2026 Week 17"
author:
  - name: "Erick K Yegon,PhD"
    email: "keyegon@gmail.com"
    orcid: "https://orcid.org/0000-0002-7055-4848"
date: today
date-format: "MMMM D, YYYY"

abstract: |
  This report provides an end-to-end analysis of US agricultural import tariffs
  sourced from the USITC Tariff Database (1997–2025). Covering all 24 HTS chapters
  in the agricultural schedule, the analysis examines Most Favored Nation (MFN) rate
  trajectories, preferential savings under NAFTA/USMCA and bilateral agreements,
  tariff-line volatility, and the structural composition of rate types. Eleven
  publication-quality visualisations, two crosstalk-linked interactive widgets,
  seven statistical tables, and formal nonparametric tests support seven key
  empirical findings.

keywords:
  - tariffs
  - trade policy
  - USITC
  - agricultural trade
  - NAFTA
  - USMCA
  - HTS

# ── Format targets ────────────────────────────────────────────
format:
  html:
    theme:
      light: [flatly, custom.scss]
      dark:  [darkly,  custom.scss]
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-link: true
    code-summary: "Show / hide code"
    code-overflow: scroll
    fig-cap-location: bottom
    tbl-cap-location: top
    reference-location: margin
    footnotes-hover: true
    embed-resources: true
    smooth-scroll: true
    highlight-style: kate
    html-math-method: katex
    include-in-header:
      text: |
        <style>
        /* Reading progress bar */
        #reading-progress {
          position: fixed; top: 0; left: 0; height: 3px;
          background: #378ADD; z-index: 9999; width: 0%;
          transition: width 0.08s linear;
        }
        h1 { border-bottom: 2px solid #E6F1FB; padding-bottom: 6px; }
        h2 { border-left: 4px solid #378ADD; padding-left: 10px; }
        .metric-row { display: flex; gap: 16px; margin: 20px 0; flex-wrap: wrap; }
        .metric-card {
          flex: 1; min-width: 140px; background: #F1EFE8;
          border-radius: 8px; padding: 14px 18px; text-align: center;
        }
        .metric-card .val { font-size: 26px; font-weight: 700;
                            color: #185FA5; display: block; }
        .metric-card .lbl { font-size: 11px; color: #5F5E5A;
                            text-transform: uppercase; letter-spacing: .04em; }
        @media print {
          #reading-progress, .sidebar { display: none !important; }
          body { font-size: 12px; }
        }
        </style>
        <div id="reading-progress"></div>
        <script>
        document.addEventListener("scroll", function() {
          var s = document.documentElement;
          var pct = (s.scrollTop / (s.scrollHeight - s.clientHeight)) * 100;
          document.getElementById("reading-progress").style.width = pct + "%";
        });
        </script>
  pdf:
    documentclass: article
    papersize: letter
    geometry: [top=1in, bottom=1in, left=1in, right=1in]
    toc: true
    number-sections: true
    colorlinks: true
    linkcolor: "#185FA5"
    fig-pos: "H"
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{float}
        \usepackage{hyperref}
  revealjs:
    theme: moon
    slide-number: c/t
    chalkboard: true
    preview-links: true
    transition: fade
    progress: true
    footer: "US Agricultural Tariffs · USITC · TidyTuesday 2026 W17"
    logo: ""
    controls: true
    auto-animate: true

# ── Execution ─────────────────────────────────────────────────
execute:
  cache: true
  freeze: auto
  warning: false
  message: false
  echo: true

# ── Cross-references ──────────────────────────────────────────
crossref:
  fig-title: "Figure"
  tbl-title: "Table"
  fig-prefix: "Fig."
  tbl-prefix: "Tbl."
  eq-prefix: "Eq."
  sec-prefix: "§"

# ── Bibliography ──────────────────────────────────────────────
# bibliography: references.bib
# csl: apa.csl
---

```{r}
#| label: setup
#| include: false
#| cache: false

knitr::opts_chunk$set(
  fig.align  = "center",
  out.width  = "100%",
  dpi        = 180,
  comment    = "#>"
)
```

```{r}
#| label: packages
#| cache: false

pkgs <- c(
  "tidyverse", "lubridate", "scales", "ggtext", "ggrepel",
  "patchwork", "gt", "gtExtras", "glue", "ggridges",
  "RColorBrewer", "forcats", "stringr", "janitor", "knitr",
  "kableExtra", "DT", "plotly", "crosstalk",
  "broom", "Kendall", "dunn.test"
)
installed  <- rownames(installed.packages())
to_install <- pkgs[!pkgs %in% installed]
if (length(to_install)) install.packages(to_install, quiet = TRUE)

suppressPackageStartupMessages({
  library(tidyverse); library(lubridate); library(scales)
  library(ggtext);    library(ggrepel);   library(patchwork)
  library(gt);        library(gtExtras);  library(glue)
  library(ggridges);  library(RColorBrewer); library(forcats)
  library(stringr);   library(janitor);   library(knitr)
  library(kableExtra);library(DT);        library(plotly)
  library(crosstalk); library(broom);     library(Kendall)
  library(dunn.test)
})
```

```{r}
#| label: theme-palette
#| cache: false

theme_tariff <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      text              = element_text(colour = "#2C2C2A"),
      plot.title        = element_markdown(size = rel(1.35), face = "bold",
                                           margin = margin(b = 6)),
      plot.subtitle     = element_markdown(size = rel(0.95), colour = "#5F5E5A",
                                           lineheight = 1.4, margin = margin(b = 12)),
      plot.caption      = element_markdown(size = rel(0.72), colour = "#888780",
                                           hjust = 0, margin = margin(t = 10)),
      plot.background   = element_rect(fill = "#FAFAF8", colour = NA),
      panel.background  = element_rect(fill = "#FAFAF8", colour = NA),
      panel.grid.major  = element_line(colour = "#D3D1C7", linewidth = 0.35),
      panel.grid.minor  = element_blank(),
      axis.title        = element_text(size = rel(0.85), colour = "#5F5E5A"),
      axis.text         = element_text(size = rel(0.80), colour = "#5F5E5A"),
      axis.ticks        = element_blank(),
      strip.text        = element_markdown(size = rel(0.85), face = "bold",
                                           colour = "#3C3489"),
      strip.background  = element_rect(fill = "#EEEDFE", colour = NA),
      legend.position   = "bottom",
      legend.title      = element_text(size = rel(0.82), colour = "#5F5E5A"),
      legend.text       = element_text(size = rel(0.78)),
      plot.margin       = margin(14, 14, 10, 14)
    )
}
theme_set(theme_tariff())

pal_section <- c(
  "Live Animals & Animal Products" = "#D55E00",
  "Vegetable Products"             = "#009E73",
  "Animal/Vegetable Fats & Oils"   = "#CC79A7",
  "Prepared Foodstuffs"            = "#0072B2"
)

pal_agree <- c(
  "MFN"         = "#534AB7",
  "NAFTA/USMCA" = "#1D9E75",
  "Bilateral"   = "#D85A30",
  "Other GSP"   = "#BA7517"
)
```

## Introduction {#sec-intro}

::: {.callout-note icon="true"}
## About this dataset
This analysis uses the **USITC Tariff Database** (TidyTuesday 2026, Week 17), covering all agricultural and food product tariff lines in HTS Chapters 1–24, with annual data from 1997 to 2025. The database includes Most Favored Nation (MFN) rates and preferential rates under NAFTA/USMCA, bilateral agreements, and the Generalised System of Preferences (GSP).
:::

**Key questions explored:**

1. How have MFN tariff rates evolved across agricultural product categories?
2. How much do NAFTA/USMCA and bilateral agreements reduce rates vs MFN?
3. Which tariff lines have been most volatile over time?
4. How did the NAFTA → USMCA transition (2020) manifest in the rate data?
5. Is there a relationship between MFN rate levels and preferential agreement coverage?
6. Are cross-section rate differences statistically significant?
7. Is there a measurable monotonic trend in MFN rates over time?

---

## Data Ingestion & Validation {#sec-data}

```{r}
#| label: load-data
#| cache: true

base_url <- paste0(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/",
  "main/data/2026/2026-04-28/"
)

safe_read <- function(file) {
  readr::read_csv(paste0(base_url, file),
                  show_col_types = FALSE, progress = FALSE)
}

agreements          <- safe_read("agreements.csv")
quantity_codes      <- safe_read("quantity_codes.csv")
tariff_agricultural <- safe_read("tariff_agricultural.csv")
tariff_codes        <- safe_read("tariff_codes.csv")
```

```{r}
#| label: tbl-datasets
#| tbl-cap: "Dataset dimensions after ingestion"

tibble(
  Dataset = c("tariff_agricultural","agreements","quantity_codes","tariff_codes"),
  Rows    = c(nrow(tariff_agricultural), nrow(agreements),
              nrow(quantity_codes),      nrow(tariff_codes)),
  Columns = c(ncol(tariff_agricultural), ncol(agreements),
              ncol(quantity_codes),      ncol(tariff_codes))
) |>
  gt() |>
  fmt_integer(columns = c(Rows, Columns)) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

```{r}
#| label: metric-cards
#| echo: false
#| results: asis

n_hts   <- n_distinct(tariff_agricultural$hts8)
n_agree <- n_distinct(tariff_agricultural$agreement)
yr_min  <- year(min(as.Date(tariff_agricultural$begin_effective_date), na.rm = TRUE))
yr_max  <- year(max(as.Date(tariff_agricultural$end_effective_date),   na.rm = TRUE))
n_rows  <- scales::comma(nrow(tariff_agricultural))

cat(glue('
<div class="metric-row">
  <div class="metric-card">
    <span class="val">{n_rows}</span>
    <span class="lbl">Total records</span>
  </div>
  <div class="metric-card">
    <span class="val">{scales::comma(n_hts)}</span>
    <span class="lbl">HTS-8 codes</span>
  </div>
  <div class="metric-card">
    <span class="val">{n_agree}</span>
    <span class="lbl">Agreement codes</span>
  </div>
  <div class="metric-card">
    <span class="val">{yr_min}–{yr_max}</span>
    <span class="lbl">Year span</span>
  </div>
  <div class="metric-card">
    <span class="val">24</span>
    <span class="lbl">HTS chapters</span>
  </div>
</div>
'))
```

::: {.column-margin}
**Quick facts**

Distinct HTS-8 codes: `r n_distinct(tariff_agricultural$hts8)`

Distinct agreements: `r n_distinct(tariff_agricultural$agreement)`

Date range: `r yr_min`–`r yr_max`
:::

---

## Data Cleaning & Feature Engineering {#sec-cleaning}

```{r}
#| label: cleaning

tariff_clean <- tariff_agricultural |>
  janitor::clean_names() |>
  mutate(
    begin_date  = as.Date(begin_effective_date),
    end_date    = as.Date(end_effective_date),
    begin_year  = year(begin_date),
    end_year    = year(end_date),
    ad_val_pct  = case_when(
      ad_val_rate >= 1    ~ ad_val_rate,
      !is.na(ad_val_rate) ~ ad_val_rate * 100,
      TRUE                ~ NA_real_
    ),
    hts_chapter = as.integer(str_sub(as.character(hts8), 1, 2)),
    section = case_when(
      hts_chapter <= 5  ~ "Live Animals & Animal Products",
      hts_chapter <= 14 ~ "Vegetable Products",
      hts_chapter <= 15 ~ "Animal/Vegetable Fats & Oils",
      hts_chapter <= 24 ~ "Prepared Foodstuffs",
      TRUE              ~ "Other"
    ),
    section = factor(section, levels = names(pal_section))
  )

YEARS <- 2000:2024

annual_panel <- tariff_clean |>
  distinct(hts8, agreement) |>
  crossing(year = YEARS) |>
  left_join(
    tariff_clean |>
      select(hts8, agreement, begin_year, end_year,
             ad_val_pct, specific_rate, hts_chapter, section),
    join_by(hts8, agreement, year >= begin_year, year <= end_year)
  ) |>
  filter(!is.na(ad_val_pct), ad_val_pct >= 0, ad_val_pct < 500) |>
  distinct(hts8, agreement, year, .keep_all = TRUE) |>
  left_join(agreements |> janitor::clean_names(), by = "agreement") |>
  mutate(
    agree_group = case_when(
      str_detect(agreement_full,
                 regex("most favored|MFN", ignore_case = TRUE))            ~ "MFN",
      str_detect(agreement_full,
                 regex("NAFTA|USMCA|Canada|Mexico", ignore_case = TRUE))   ~ "NAFTA/USMCA",
      str_detect(agreement_full,
                 regex("bilateral|Australia|Chile|Korea|Colombia|Peru|Singapore",
                       ignore_case = TRUE))                                 ~ "Bilateral",
      TRUE                                                                  ~ "Other GSP"
    ),
    agree_group = factor(agree_group, levels = names(pal_agree))
  )

mfn_panel <- annual_panel |>
  filter(agree_group == "MFN") |>
  select(hts8, year, section, hts_chapter, mfn_rate = ad_val_pct)
```

::: {.callout-tip}
## Pipeline design note
The annual panel uses an **inequality join** (`join_by(year >= begin_year, year <= end_year)`) to correctly assign whichever tariff rate was legally in force for every HTS × agreement × year cell, avoiding rate-smearing across policy change dates.
:::

Annual panel rows: `r scales::comma(nrow(annual_panel))` · MFN panel rows: `r scales::comma(nrow(mfn_panel))`

---

## Exploratory Data Analysis {#sec-eda}

### Descriptive statistics by section {#sec-desc-stats}

```{r}
#| label: tbl-section-stats
#| tbl-cap: "MFN rate summary statistics by HTS section (2000–2024)"

section_stats <- mfn_panel |>
  group_by(section) |>
  summarise(
    `Tariff lines` = n_distinct(hts8),
    `Obs. (n)`     = n(),
    `Mean rate`    = mean(mfn_rate, na.rm = TRUE),
    `Median`       = median(mfn_rate, na.rm = TRUE),
    `SD`           = sd(mfn_rate, na.rm = TRUE),
    `Min`          = min(mfn_rate, na.rm = TRUE),
    `Max`          = max(mfn_rate, na.rm = TRUE),
    `% zero-rated` = mean(mfn_rate == 0, na.rm = TRUE) * 100,
    .groups = "drop"
  )

section_stats |>
  gt() |>
  tab_header(
    title    = md("**MFN Rate Summary Statistics by HTS Section (2000–2024)**"),
    subtitle = md("*Ad valorem rates (%). Source: USITC Tariff Database.*")
  ) |>
  cols_label(section = "Section") |>
  fmt_number(columns = c(`Mean rate`, Median, SD, Min, Max),
             decimals = 2, suffix = "%") |>
  fmt_number(columns = `% zero-rated`, decimals = 1, suffix = "%") |>
  fmt_integer(columns = c(`Tariff lines`, `Obs. (n)`)) |>
  data_color(columns = `Mean rate`,    palette = c("#E6F1FB","#185FA5")) |>
  data_color(columns = `% zero-rated`, palette = c("#185FA5","#E6F1FB")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13),
              column_labels.border.top.color = "#534AB7",
              column_labels.border.top.width = px(2))
```

### Agreement group coverage {#sec-agree-coverage}

```{r}
#| label: tbl-agreement-coverage
#| tbl-cap: "Coverage and average rates by trade agreement group"

agreement_summary <- annual_panel |>
  group_by(agree_group) |>
  summarise(
    `Distinct products` = n_distinct(hts8),
    `Product-years`     = n(),
    `Mean rate`         = mean(ad_val_pct, na.rm = TRUE),
    `Median rate`       = median(ad_val_pct, na.rm = TRUE),
    .groups = "drop"
  )

agreement_summary |>
  gt() |>
  cols_label(agree_group = "Agreement group") |>
  fmt_number(columns = c(`Mean rate`, `Median rate`), decimals = 2, suffix = "%") |>
  fmt_integer(columns = c(`Distinct products`, `Product-years`)) |>
  data_color(columns = `Mean rate`, palette = c("#E6F1FB","#185FA5")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### MFN rate vs preferential saving — OLS regression {#sec-regression}

```{r}
#| label: fig-regression
#| fig-cap: "OLS regression: does a higher MFN rate predict a larger NAFTA/USMCA saving? Each point = one HTS chapter."
#| fig-width: 10
#| fig-height: 6

reg_data <- annual_panel |>
  filter(agree_group %in% c("MFN","NAFTA/USMCA")) |>
  group_by(hts_chapter, section, agree_group) |>
  summarise(avg = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = agree_group, values_from = avg) |>
  janitor::clean_names() |>
  filter(!is.na(mfn), !is.na(nafta_usmca)) |>
  mutate(saving = mfn - nafta_usmca)

model <- lm(saving ~ mfn, data = reg_data)
r2    <- glance(model)$r.squared
slope <- tidy(model)$estimate[2]

ggplot(reg_data, aes(x = mfn, y = saving, colour = section)) +
  geom_smooth(method = "lm", se = TRUE, colour = "#534AB7",
              fill = "#EEEDFE", linewidth = 0.8, alpha = 0.25) +
  geom_point(size = 4, alpha = 0.85) +
  geom_text_repel(aes(label = glue("Ch.{hts_chapter}")),
                  size = 2.8, max.overlaps = 20,
                  segment.colour = "#B4B2A9") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%")) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  annotate("richtext", x = -Inf, y = Inf, hjust = -0.05, vjust = 1.3,
           label = glue("R² = {round(r2,3)} · slope = {round(slope,2)}"),
           fill = NA, label.colour = NA, size = 3.5, colour = "#534AB7") +
  labs(
    title    = "Higher MFN rates predict larger NAFTA/USMCA tariff savings",
    subtitle = "OLS regression across HTS chapters. Shaded band = 95% CI.",
    x = "Mean MFN rate", y = "Mean NAFTA/USMCA saving (pp)",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  )
```

```{r}
#| label: tbl-regression
#| tbl-cap: "OLS regression output: NAFTA/USMCA saving ~ MFN rate"

bind_rows(
  tidy(model) |> mutate(type = "Coefficients"),
  glance(model) |>
    select(r.squared, adj.r.squared, p.value, nobs) |>
    pivot_longer(everything(), names_to = "term", values_to = "estimate") |>
    mutate(type = "Model fit", std.error = NA_real_,
           statistic = NA_real_, p.value = NA_real_)
) |>
  gt() |>
  tab_header(title    = md("**OLS: NAFTA/USMCA saving ~ MFN rate**"),
             subtitle = md("*Unit of observation: HTS chapter*")) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  tab_row_group(label = "Coefficients", rows = type == "Coefficients") |>
  tab_row_group(label = "Model fit",    rows = type == "Model fit") |>
  cols_hide(type) |>
  tab_options(table.font.size = px(13))
```

### Tariff dispersion — Lorenz / Gini analysis {#sec-gini}

```{r}
#| label: fig-gini
#| fig-cap: "Lorenz curves for MFN tariff distribution by section (latest year). Convexity = inequality of tariff burden."
#| fig-width: 9
#| fig-height: 6

gini_fn <- function(x) {
  x <- sort(x[x > 0 & !is.na(x)])
  n <- length(x)
  if (n < 2) return(NA_real_)
  (2 * sum(seq_len(n) * x) / (n * sum(x))) - (n + 1) / n
}

lorenz_df <- mfn_panel |>
  filter(year == max(year)) |>
  group_by(section) |>
  filter(mfn_rate > 0) |>
  arrange(mfn_rate) |>
  mutate(
    cum_pop  = row_number() / n(),
    cum_rate = cumsum(mfn_rate) / sum(mfn_rate)
  ) |>
  ungroup()

gini_vals <- mfn_panel |>
  filter(year == max(year)) |>
  group_by(section) |>
  summarise(gini = gini_fn(mfn_rate), .groups = "drop") |>
  mutate(label = sprintf("Gini = %.3f", gini),
         x = 0.18, y = 0.80 - (as.integer(section) - 1) * 0.09)

ggplot(lorenz_df, aes(x = cum_pop, y = cum_rate, colour = section)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "#888780", linewidth = 0.6) +
  geom_line(linewidth = 1.2) +
  geom_label(data = gini_vals,
             aes(x = x, y = y,
                 label = paste0(str_wrap(as.character(section), 20), "\n", label),
                 colour = section),
             size = 2.5, fill = "#FAFAF8", label.size = 0.3,
             show.legend = FALSE) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_x_continuous(labels = label_percent()) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    title    = "Lorenz curves: inequality of MFN tariff rates by section",
    subtitle = glue("Year: {max(mfn_panel$year)}. Dashed = perfect equality."),
    x = "Cumulative share of tariff lines",
    y = "Cumulative share of total tariff rate",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  )
```

---

## Visualisations {#sec-viz}

### MFN rate trajectories {#sec-fig-trajectories}

```{r}
#| label: fig-trajectories
#| fig-cap: "Mean MFN ad valorem rate by HTS section (2000–2024). Ribbon = ±1 SD. Grey = all sections for context."
#| fig-width: 11
#| fig-height: 7

section_yr <- mfn_panel |>
  group_by(section, year) |>
  summarise(
    mean_rate  = mean(mfn_rate, na.rm = TRUE),
    sd_rate    = sd(mfn_rate, na.rm = TRUE),
    n_products = n_distinct(hts8),
    .groups = "drop"
  ) |>
  filter(n_products >= 5) |>
  mutate(lo = pmax(mean_rate - sd_rate, 0), hi = mean_rate + sd_rate)

bg_lines <- section_yr |> rename(bg_section = section)

ggplot() +
  geom_line(data = bg_lines,
            aes(x = year, y = mean_rate, group = bg_section),
            colour = "#D3D1C7", linewidth = 0.7, alpha = 0.6) +
  geom_ribbon(data = section_yr,
              aes(x = year, ymin = lo, ymax = hi, fill = section),
              alpha = 0.18) +
  geom_line(data = section_yr,
            aes(x = year, y = mean_rate, colour = section),
            linewidth = 1.3) +
  geom_point(data = section_yr,
             aes(x = year, y = mean_rate, colour = section),
             size = 1.6, shape = 21, fill = "white", stroke = 1.0) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_fill_manual(values = pal_section, guide = "none") +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  scale_y_continuous(labels = label_number(suffix = "%"), limits = c(0, NA)) +
  facet_wrap(~section, ncol = 2, scales = "free_y") +
  labs(
    title    = "MFN tariff trends by agricultural section (2000–2024)",
    subtitle = "Mean rate ± 1 SD (ribbon). Grey = all sections for comparison.",
    x = NULL, y = "Average MFN rate",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  )
```

::: {.callout-note appearance="minimal"}
Live Animals & Animal Products sustains the highest MFN average throughout the period. Vegetable Products shows a modest downward trend through the 2010s.
:::

### Year-over-year rate change heatmap {#sec-fig-yoy}

```{r}
#| label: fig-yoy
#| fig-cap: "Year-over-year change in mean MFN rate per section. Red = increase, blue = decrease. Labels shown for |Δ| ≥ 0.3 pp."
#| fig-height: 5.5

yoy_data <- mfn_panel |>
  group_by(section, year) |>
  summarise(mean_rate = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  arrange(section, year) |>
  group_by(section) |>
  mutate(yoy_change = mean_rate - lag(mean_rate)) |>
  ungroup() |>
  filter(!is.na(yoy_change))

ggplot(yoy_data, aes(x = year, y = fct_rev(section), fill = yoy_change)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = ifelse(abs(yoy_change) >= 0.3,
                               sprintf("%+.1f", yoy_change), "")),
            size = 2.5, colour = "white", fontface = "bold") +
  scale_fill_gradient2(
    low = "#185FA5", mid = "#F1EFE8", high = "#D85A30", midpoint = 0,
    name = "YoY Δ (pp)",
    guide = guide_colourbar(barwidth = 10, barheight = 0.5,
                            title.position = "top", title.hjust = 0.5)
  ) +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  labs(
    title    = "Year-over-year MFN rate changes by section",
    subtitle = "Red = increase · Blue = decrease · Labels for |Δ| ≥ 0.3 pp.",
    x = NULL, y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_markdown(size = 9))
```

### Agreement coverage heatmap {#sec-fig-heatmap}

```{r}
#| label: fig-heatmap
#| fig-cap: "Average ad valorem rate by HTS chapter and agreement group. Darker = higher rates."
#| fig-width: 10
#| fig-height: 13

heatmap_data <- annual_panel |>
  mutate(chapter_label = glue("Ch. {str_pad(hts_chapter, 2, 'left', '0')}")) |>
  group_by(section, hts_chapter, chapter_label, agree_group) |>
  summarise(avg_rate = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  filter(agree_group != "Other GSP") |>
  mutate(
    agree_group   = factor(agree_group, levels = c("MFN","NAFTA/USMCA","Bilateral")),
    chapter_label = fct_reorder(chapter_label, hts_chapter)
  )

ggplot(heatmap_data,
       aes(x = agree_group, y = fct_rev(chapter_label), fill = avg_rate)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(avg_rate >= 1, sprintf("%.1f%%", avg_rate), "")),
            size = 2.5, colour = "white", fontface = "bold") +
  scale_fill_gradientn(
    colours  = c("#E6F1FB","#378ADD","#185FA5","#042C53"),
    na.value = "#F1EFE8", name = "Avg. rate",
    labels   = label_number(suffix = "%"),
    guide    = guide_colourbar(barwidth = 10, barheight = 0.5,
                               title.position = "top", title.hjust = 0.5)
  ) +
  facet_grid(section ~ ., scales = "free_y", space = "free_y") +
  labs(
    title    = "Tariff rate by HTS chapter and trade agreement",
    subtitle = "Each cell = mean ad valorem rate for that chapter–agreement pair.",
    x = "Agreement group", y = "HTS chapter",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(panel.grid   = element_blank(),
        axis.text.y  = element_text(size = 7),
        strip.text.y = element_text(angle = 0, hjust = 0, size = 7))
```

### MFN vs NAFTA/USMCA dumbbell {#sec-fig-dumbbell}

```{r}
#| label: fig-dumbbell
#| fig-cap: "MFN (filled) vs NAFTA/USMCA preferential (open) rates per HTS chapter. Segment = pp saving."
#| fig-width: 10
#| fig-height: 8

dumbbell_data <- annual_panel |>
  filter(agree_group %in% c("MFN","NAFTA/USMCA")) |>
  group_by(hts_chapter, section, agree_group) |>
  summarise(avg_rate = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = agree_group, values_from = avg_rate) |>
  clean_names() |>
  filter(!is.na(mfn), !is.na(nafta_usmca)) |>
  mutate(
    saving        = mfn - nafta_usmca,
    chapter_label = glue("Ch. {str_pad(hts_chapter, 2, 'left', '0')}"),
    chapter_label = fct_reorder(chapter_label, saving)
  ) |>
  filter(saving > 0)

ggplot(dumbbell_data) +
  geom_segment(aes(x = nafta_usmca, xend = mfn,
                   y = chapter_label, yend = chapter_label, colour = section),
               linewidth = 1.8, alpha = 0.40) +
  geom_point(aes(x = nafta_usmca, y = chapter_label, colour = section),
             size = 3.8, shape = 21, fill = "white", stroke = 1.2) +
  geom_point(aes(x = mfn, y = chapter_label, colour = section), size = 3.8) +
  geom_text(aes(x = mfn + 0.25, y = chapter_label,
                label = sprintf("−%.1f pp", saving)),
            hjust = 0, size = 2.9, colour = "#5F5E5A") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0.01, 0.16))) +
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(
    title    = "NAFTA/USMCA preferential savings vs MFN by chapter",
    subtitle = "Open = NAFTA/USMCA rate · Filled = MFN · Label = pp reduction.",
    x = "Average ad valorem rate", y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(panel.grid.major.y = element_blank())
```

### Tariff volatility — top 30 lines {#sec-fig-volatility}

```{r}
#| label: fig-volatility
#| fig-cap: "Top 30 MFN tariff lines by rate standard deviation, 2000–2024."
#| fig-width: 12
#| fig-height: 10

volatility <- mfn_panel |>
  group_by(hts8, section) |>
  filter(n() >= 5) |>
  summarise(rate_sd  = sd(mfn_rate, na.rm = TRUE),
            mean_mfn = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  filter(mean_mfn > 0)

top_volatile <- volatility |>
  slice_max(rate_sd, n = 30) |>
  left_join(
    {
      tc       <- tariff_codes |> janitor::clean_names()
      desc_col <- names(tc)[str_detect(names(tc), "brief|description|desc")][1]
      if (!is.na(desc_col)) {
        tc |> distinct(hts8, .keep_all = TRUE) |>
          select(hts8, description = all_of(desc_col))
      } else {
        tc |> distinct(hts8) |> mutate(description = NA_character_)
      }
    },
    by = "hts8"
  ) |>
  mutate(label = coalesce(str_trunc(description, 45), as.character(hts8)),
         label = fct_reorder(label, rate_sd))

ggplot(top_volatile, aes(x = rate_sd, y = label, colour = section)) +
  geom_segment(aes(x = 0, xend = rate_sd, y = label, yend = label),
               linewidth = 0.9, alpha = 0.50) +
  geom_point(size = 3.6) +
  geom_text(aes(label = sprintf("%.1f%%", rate_sd)),
            hjust = -0.3, size = 2.7, colour = "#5F5E5A") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0, 0.18))) +
  labs(
    title    = "Top 30 most volatile MFN tariff lines",
    subtitle = "Ranked by standard deviation of annual MFN rate, 2000–2024.",
    x = "Rate SD (pp)", y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y        = element_text(size = 7.5))
```

### NAFTA → USMCA transition {#sec-fig-nafta}

```{r}
#| label: fig-nafta
#| fig-cap: "NAFTA/USMCA preferential rate trajectories by section. Dashed = USMCA effective date (2020)."
#| fig-width: 11
#| fig-height: 6

nafta_section <- annual_panel |>
  filter(agree_group == "NAFTA/USMCA") |>
  group_by(year, section) |>
  summarise(
    mean_rate = mean(ad_val_pct, na.rm = TRUE),
    lo        = quantile(ad_val_pct, 0.25, na.rm = TRUE),
    hi        = quantile(ad_val_pct, 0.75, na.rm = TRUE),
    .groups   = "drop"
  )

ggplot(nafta_section, aes(x = year, colour = section, fill = section)) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             colour = "#534AB7", linewidth = 0.7, alpha = 0.7) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.14, linewidth = 0) +
  geom_line(aes(y = mean_rate), linewidth = 1.2) +
  geom_point(aes(y = mean_rate), size = 1.7, shape = 21, fill = "white") +
  annotate("richtext", x = 2020.4, y = Inf, hjust = 0, vjust = 1.4,
           size = 3.2, label = "**USMCA** replaces NAFTA",
           fill = NA, label.colour = NA, colour = "#534AB7") +
  scale_colour_manual(values = pal_section, name = NULL) +
  scale_fill_manual(values = pal_section, name = NULL) +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  scale_y_continuous(labels = label_number(suffix = "%"), limits = c(0, NA)) +
  labs(
    title    = "NAFTA → USMCA: preferential tariff rate evolution",
    subtitle = "Mean (line) ± IQR (ribbon). USMCA effective July 2020.",
    x = NULL, y = "Average preferential rate",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(legend.position = "top")
```

### Biggest rate movers {#sec-fig-movers}

```{r}
#| label: fig-movers
#| fig-cap: "Top 15 MFN rate increases and decreases (earliest → latest observation per tariff line)."
#| fig-width: 13
#| fig-height: 10

big_movers <- mfn_panel |>
  group_by(hts8, section) |>
  filter(n_distinct(year) >= 10) |>
  summarise(
    rate_start = mean(mfn_rate[year == min(year)], na.rm = TRUE),
    rate_end   = mean(mfn_rate[year == max(year)], na.rm = TRUE),
    change     = rate_end - rate_start,
    .groups    = "drop"
  ) |>
  filter(!is.na(change)) |>
  mutate(direction = if_else(change > 0, "Increase", "Decrease")) |>
  group_by(direction) |>
  slice_max(abs(change), n = 15) |>
  ungroup() |>
  left_join(
    {
      tc       <- tariff_codes |> janitor::clean_names()
      desc_col <- names(tc)[str_detect(names(tc), "brief|description|desc")][1]
      if (!is.na(desc_col)) {
        tc |> distinct(hts8, .keep_all = TRUE) |>
          select(hts8, description = all_of(desc_col))
      } else {
        tc |> distinct(hts8) |> mutate(description = NA_character_)
      }
    },
    by = "hts8"
  ) |>
  mutate(label = coalesce(str_trunc(description, 40), as.character(hts8)),
         label = fct_reorder(label, change))

ggplot(big_movers, aes(x = change, y = label, fill = section)) +
  geom_col(alpha = 0.82, width = 0.72) +
  geom_vline(xintercept = 0, colour = "#2C2C2A", linewidth = 0.65) +
  geom_text(aes(label = sprintf("%+.1f pp", change),
                hjust = if_else(change > 0, -0.15, 1.15)),
            size = 2.8, colour = "#2C2C2A") +
  scale_fill_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = " pp", style_positive = "plus"),
                     expand = expansion(mult = 0.20)) +
  labs(
    title    = "Biggest MFN tariff changes: earliest → latest observation",
    subtitle = "Top 15 increases and decreases in percentage-point terms.",
    x = "Change (pp)", y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y        = element_markdown(size = 7.2))
```

### Agreements vs MFN rate — bubble chart {#sec-fig-bubble}

```{r}
#| label: fig-bubble
#| fig-cap: "HTS chapters: avg preferential agreement groups vs mean MFN rate. Bubble size = tariff lines."
#| fig-width: 10
#| fig-height: 7

bubble_chapter <- annual_panel |>
  mutate(hts_chapter = as.integer(str_sub(as.character(hts8), 1, 2))) |>
  group_by(hts8, hts_chapter, section) |>
  summarise(
    n_agreements = n_distinct(agree_group[agree_group != "MFN"]),
    mean_mfn     = mean(ad_val_pct[agree_group == "MFN"], na.rm = TRUE),
    .groups      = "drop"
  ) |>
  filter(!is.na(mean_mfn), mean_mfn > 0) |>
  group_by(hts_chapter, section) |>
  summarise(
    n_agreements = mean(n_agreements, na.rm = TRUE),
    mean_mfn     = mean(mean_mfn, na.rm = TRUE),
    n_products   = n(),
    .groups      = "drop"
  ) |>
  mutate(chapter_label = glue("Ch.{hts_chapter}"))

ggplot(bubble_chapter,
       aes(x = n_agreements, y = mean_mfn,
           size = n_products, colour = section)) +
  geom_point(alpha = 0.72) +
  geom_text_repel(aes(label = chapter_label), size = 2.7, max.overlaps = 20,
                  segment.colour = "#B4B2A9", segment.linewidth = 0.4) +
  scale_size_area(max_size = 14, name = "# products") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(breaks = 0:5, minor_breaks = NULL) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  guides(size   = guide_legend(override.aes = list(colour = "#888780")),
         colour = guide_legend(override.aes = list(size = 4))) +
  labs(
    title    = "Do high-tariff chapters attract more preferential agreements?",
    subtitle = "Bubble = HTS chapter. Size = tariff lines. X = avg. agreement groups.",
    x = "Average number of preferential agreement groups",
    y = "Mean MFN tariff rate",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  )
```

### Tariff type composition {#sec-fig-composition}

```{r}
#| label: fig-composition
#| fig-cap: "Share of MFN tariff lines by rate structure type over time and section."
#| fig-width: 11
#| fig-height: 7

type_pal <- c("Ad valorem only" = "#378ADD", "Specific only" = "#D85A30",
              "Compound (both)" = "#BA7517", "Zero / unknown" = "#D3D1C7")

composition_data <- annual_panel |>
  filter(agree_group == "MFN") |>
  mutate(
    rate_type = case_when(
      !is.na(specific_rate) & specific_rate > 0 &
        !is.na(ad_val_pct)   & ad_val_pct   > 0  ~ "Compound (both)",
      !is.na(specific_rate) & specific_rate > 0  ~ "Specific only",
      !is.na(ad_val_pct)   & ad_val_pct   > 0  ~ "Ad valorem only",
      TRUE                                       ~ "Zero / unknown"
    )
  ) |>
  count(year, section, rate_type) |>
  group_by(year, section) |>
  mutate(share = n / sum(n)) |>
  ungroup()

ggplot(composition_data, aes(x = year, y = share, fill = rate_type)) +
  geom_area(alpha = 0.85, position = "stack") +
  scale_fill_manual(values = type_pal, name = "Tariff type") +
  scale_x_continuous(breaks = seq(2000, 2024, 6)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  facet_wrap(~section, ncol = 2) +
  labs(
    title    = "Composition of MFN tariff rate structures by section",
    subtitle = "Share of lines with ad valorem, specific, or compound rates.",
    x = NULL, y = "Share of tariff lines",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(legend.position = "bottom")
```

### Rate distribution ridgelines {#sec-fig-ridges}

```{r}
#| label: fig-ridges
#| fig-cap: "MFN ad valorem rate distributions every four years, by section. X-axis capped at 60%."
#| fig-width: 11
#| fig-height: 8

mfn_panel |>
  filter(year %in% seq(2000, 2024, 4), mfn_rate > 0, mfn_rate <= 100) |>
  mutate(year_f = factor(year)) |>
  ggplot(aes(x = mfn_rate, y = year_f, fill = section)) +
  geom_density_ridges(alpha = 0.65, scale = 1.8,
                      rel_min_height = 0.01, bandwidth = 1.5) +
  scale_fill_manual(values = pal_section, name = NULL) +
  scale_x_continuous(labels = label_number(suffix = "%"), limits = c(0, 60)) +
  facet_wrap(~section, ncol = 2) +
  labs(
    title    = "Distribution of MFN tariff rates over time (every 4 years)",
    subtitle = "X-axis capped at 60% for readability.",
    x = "MFN ad valorem rate", y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 Week 17"
  ) +
  theme(legend.position = "none")
```

---

## Statistical Tests {#sec-stats}

### Kruskal-Wallis: are section rates significantly different?

```{r}
#| label: tbl-kruskal
#| tbl-cap: "Kruskal-Wallis test: MFN rates across HTS sections"

kw <- kruskal.test(mfn_rate ~ section, data = mfn_panel)

tibble(
  Test      = "Kruskal-Wallis rank-sum",
  Statistic = round(kw$statistic, 2),
  df        = kw$parameter,
  `p-value` = formatC(kw$p.value, format = "e", digits = 3),
  Result    = if_else(kw$p.value < 0.05,
                      "Sections differ significantly (p < 0.05)",
                      "No significant difference")
) |>
  gt() |>
  tab_header(title    = md("**Kruskal-Wallis test**"),
             subtitle = md("*H₀: all sections have identical rate distributions*")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### Dunn post-hoc pairwise comparisons (Bonferroni)

```{r}
#| label: tbl-dunn
#| tbl-cap: "Dunn post-hoc test — pairwise MFN rate comparisons (Bonferroni correction)"

dunn_res <- dunn.test(
  mfn_panel$mfn_rate,
  mfn_panel$section,
  method = "bonferroni",
  alpha  = 0.05,
  kw     = FALSE,
  label  = TRUE
)

tibble(
  Comparison   = dunn_res$comparisons,
  Z            = round(dunn_res$Z, 3),
  `p-adjusted` = round(dunn_res$P.adjusted, 4),
  Significant  = if_else(dunn_res$P.adjusted < 0.05, "Yes ✓", "No")
) |>
  gt() |>
  tab_header(title    = md("**Dunn post-hoc test (Bonferroni)**"),
             subtitle = md("*Pairwise section comparisons*")) |>
  tab_style(style = cell_fill(color = "#EAF3DE"),
            locations = cells_body(rows = Significant == "Yes ✓")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### Mann-Kendall monotonic trend test

```{r}
#| label: tbl-mannkendall
#| tbl-cap: "Mann-Kendall trend test by section (2000–2024)"

mk_results <- mfn_panel |>
  group_by(section, year) |>
  summarise(mean_rate = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  group_by(section) |>
  summarise(
    tau     = MannKendall(mean_rate)$tau[[1]],
    p_value = MannKendall(mean_rate)$sl[[1]],
    trend   = case_when(
      p_value >= 0.05           ~ "No trend",
      tau < 0 & p_value < 0.05 ~ "Significant decrease ↓",
      tau > 0 & p_value < 0.05 ~ "Significant increase ↑"
    ),
    .groups = "drop"
  )

mk_results |>
  gt() |>
  tab_header(title    = md("**Mann-Kendall trend test by section (2000–2024)**"),
             subtitle = md("*H₀: no monotonic trend in annual mean MFN rate*")) |>
  cols_label(section = "Section", tau = "Kendall τ",
             p_value = "p-value", trend = "Conclusion") |>
  fmt_number(columns = c(tau, p_value), decimals = 4) |>
  tab_style(style = cell_fill(color = "#E6F1FB"),
            locations = cells_body(rows = str_detect(trend, "decrease"))) |>
  tab_style(style = cell_fill(color = "#FAEEDA"),
            locations = cells_body(rows = str_detect(trend, "increase"))) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

---

## Interactive Exploration {#sec-interactive}

### Linked chapter explorer {#sec-crosstalk}

```{r}
#| label: crosstalk-filter

shared_chapter <- mfn_panel |>
  group_by(hts_chapter, section) |>
  summarise(
    Products  = n_distinct(hts8),
    Mean_rate = round(mean(mfn_rate, na.rm = TRUE), 2),
    Median    = round(median(mfn_rate, na.rm = TRUE), 2),
    SD        = round(sd(mfn_rate, na.rm = TRUE), 2),
    Max       = round(max(mfn_rate, na.rm = TRUE), 2),
    .groups   = "drop"
  ) |>
  rename(Chapter = hts_chapter, Section = section) |>
  SharedData$new(key = ~Chapter)

filter_select("sec_filter", "Filter by section:", shared_chapter, ~Section)
```

```{r}
#| label: tbl-crosstalk
#| tbl-cap: "MFN rate statistics by HTS chapter — select rows to highlight in scatter below"

DT::datatable(
  shared_chapter,
  rownames   = FALSE,
  extensions = "Buttons",
  options    = list(dom = "Bfrtip", buttons = c("copy","csv"),
                    pageLength = 12, scrollX = TRUE)
) |>
  DT::formatStyle(
    "Mean_rate",
    background         = DT::styleColorBar(c(0, 30), "#378ADD"),
    backgroundSize     = "100% 90%",
    backgroundRepeat   = "no-repeat",
    backgroundPosition = "center"
  )
```

```{r}
#| label: fig-crosstalk-scatter
#| fig-cap: "Linked scatter — select rows in the table above to highlight chapters here."

plot_ly(shared_chapter,
        x = ~Mean_rate, y = ~SD, size = ~Products, color = ~Section,
        colors = unname(pal_section), type = "scatter", mode = "markers",
        marker = list(opacity = 0.75, sizemode = "diameter"),
        hovertemplate = paste0(
          "<b>Chapter %{customdata}</b><br>",
          "Mean MFN: %{x:.1f}%<br>SD: %{y:.1f}%<extra></extra>"
        ),
        customdata = ~Chapter
) |>
  layout(
    title  = list(text = "Chapter mean rate vs volatility (linked to table)"),
    xaxis  = list(title = "Mean MFN rate (%)", gridcolor = "#D3D1C7"),
    yaxis  = list(title = "Rate SD (pp)", gridcolor = "#D3D1C7"),
    legend = list(orientation = "h", y = -0.2),
    paper_bgcolor = "#FAFAF8", plot_bgcolor = "#FAFAF8"
  ) |>
  highlight(on = "plotly_selected", off = "plotly_deselect",
            color = "#D85A30", persistent = TRUE)
```

### Interactive MFN trajectories with IQR bands {#sec-plotly-trajectories}

```{r}
#| label: fig-plotly-trajectories
#| fig-cap: "Interactive MFN trajectories with IQR bands — hover uses unified x tooltip."

section_yr_plotly <- mfn_panel |>
  group_by(section, year) |>
  summarise(
    mean_rate = mean(mfn_rate, na.rm = TRUE),
    lo        = quantile(mfn_rate, 0.25, na.rm = TRUE),
    hi        = quantile(mfn_rate, 0.75, na.rm = TRUE),
    n         = n_distinct(hts8),
    .groups   = "drop"
  )

p <- plot_ly()
for (sec in levels(section_yr_plotly$section)) {
  d   <- filter(section_yr_plotly, section == sec)
  col <- pal_section[sec]
  p   <- add_ribbons(p, data = d, x = ~year, ymin = ~lo, ymax = ~hi,
                     fillcolor = paste0(col, "22"), line = list(width = 0),
                     name = paste0(sec, " IQR"), legendgroup = sec,
                     showlegend = FALSE, hoverinfo = "skip")
  p   <- add_lines(p, data = d, x = ~year, y = ~mean_rate,
                   line = list(color = col, width = 2.5),
                   name = sec, legendgroup = sec,
                   hovertemplate = paste0(
                     "<b>", sec, "</b><br>Year: %{x}<br>",
                     "Mean: %{y:.2f}%<extra></extra>"))
}
p |> layout(
  title     = list(text = "MFN rate trajectories with IQR bands", font = list(size = 14)),
  xaxis     = list(title = "", tickvals = seq(2000, 2024, 4), gridcolor = "#D3D1C7"),
  yaxis     = list(title = "Average MFN rate (%)", rangemode = "tozero",
                   gridcolor = "#D3D1C7"),
  legend    = list(orientation = "h", y = -0.2),
  hovermode = "x unified",
  paper_bgcolor = "#FAFAF8", plot_bgcolor = "#FAFAF8"
)
```

---

## Composite Dashboard {#sec-dashboard}

```{r}
#| label: fig-dashboard
#| fig-cap: "Four-panel summary dashboard."
#| fig-width: 16
#| fig-height: 10
#| column: screen

p1 <- ggplot() +
  geom_line(data = bg_lines,
            aes(x = year, y = mean_rate, group = bg_section),
            colour = "#D3D1C7", linewidth = 0.5, alpha = 0.6) +
  geom_ribbon(data = section_yr,
              aes(x = year, ymin = lo, ymax = hi, fill = section), alpha = 0.15) +
  geom_line(data = section_yr,
            aes(x = year, y = mean_rate, colour = section), linewidth = 1.0) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_fill_manual(values = pal_section, guide = "none") +
  scale_x_continuous(breaks = seq(2000, 2024, 8)) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  facet_wrap(~section, ncol = 2, scales = "free_y") +
  labs(title = "MFN trajectories", x = NULL, y = "Mean rate") +
  theme(plot.title = element_markdown(size = 9.5, face = "bold"),
        strip.text = element_markdown(size = 6.5))

p3 <- ggplot(dumbbell_data) +
  geom_segment(aes(x = nafta_usmca, xend = mfn,
                   y = chapter_label, yend = chapter_label, colour = section),
               linewidth = 1.4, alpha = 0.40) +
  geom_point(aes(x = nafta_usmca, y = chapter_label, colour = section),
             size = 2.2, shape = 21, fill = "white") +
  geom_point(aes(x = mfn, y = chapter_label, colour = section), size = 2.2) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(title = "MFN vs NAFTA/USMCA", x = "Avg rate", y = NULL) +
  theme(plot.title = element_markdown(size = 9.5, face = "bold"),
        axis.text.y = element_text(size = 6),
        panel.grid.major.y = element_blank())

p5 <- ggplot(nafta_section, aes(x = year, colour = section, fill = section)) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             colour = "#534AB7", linewidth = 0.6) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.13, linewidth = 0) +
  geom_line(aes(y = mean_rate), linewidth = 1.0) +
  scale_colour_manual(values = pal_section, name = NULL) +
  scale_fill_manual(values = pal_section, name = NULL) +
  scale_x_continuous(breaks = seq(2000, 2024, 8)) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  labs(title = "NAFTA → USMCA", x = NULL, y = "Pref. rate") +
  theme(plot.title = element_markdown(size = 9.5, face = "bold"),
        legend.position = "none")

p7 <- ggplot(bubble_chapter,
             aes(x = n_agreements, y = mean_mfn,
                 size = n_products, colour = section)) +
  geom_point(alpha = 0.7) +
  geom_text_repel(aes(label = chapter_label), size = 2.1, max.overlaps = 15,
                  segment.colour = "#B4B2A9", segment.linewidth = 0.3) +
  scale_size_area(max_size = 10, guide = "none") +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_x_continuous(breaks = 0:4) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  labs(title = "Agreements vs MFN", x = "Avg agreement groups", y = "Mean MFN rate") +
  theme(plot.title = element_markdown(size = 9.5, face = "bold"))

(p1 | p3) / (p5 | p7) +
  plot_annotation(
    title    = "US Agricultural Tariff Analysis — Dashboard",
    subtitle = "USITC Tariff Database · HTS Ch. 1–24 · TidyTuesday 2026 Week 17",
    theme    = theme(
      plot.title    = element_markdown(size = 14, face = "bold", colour = "#185FA5"),
      plot.subtitle = element_markdown(size = 9, colour = "#5F5E5A")
    )
  )
```

---

## Key Findings & Conclusions {#sec-conclusions}

```{r}
#| label: tbl-findings
#| tbl-cap: "Summary of key analytical findings"

tibble(
  `#` = 1:7,
  Finding = c(
    "Live Animals & Animal Products consistently carry the highest MFN rates",
    "NAFTA/USMCA delivers substantial tariff reductions across most chapters",
    "Prepared Foodstuffs shows the most stable MFN rates over time",
    "Rate volatility is concentrated in a small subset of tariff lines",
    "The NAFTA → USMCA transition (2020) did not cause a discontinuous rate jump",
    "Chapters with higher MFN rates tend to have more preferential agreement coverage",
    "MFN rate differences across sections are statistically significant (Kruskal-Wallis)"
  ),
  Implication = c(
    "Animal products remain a sensitive sector with high import protection",
    "USMCA trade partners enjoy significant competitive advantages over MFN countries",
    "Processed food tariffs are politically stable and less subject to renegotiation",
    "Monitoring ~30 lines captures most of the system-wide tariff uncertainty",
    "USMCA largely preserved existing NAFTA preferential rates during transition",
    "High-tariff sectors generate more demand for bilateral trade deals",
    "Section-level analysis is analytically justified — sections are not exchangeable"
  )
) |>
  gt() |>
  tab_header(title    = md("**Key Findings — US Agricultural Tariff Analysis**"),
             subtitle = md("*MFN and preferential rate trends, 2000–2024*")) |>
  cols_label(`#` = "#") |>
  cols_width(`#` ~ px(28), Finding ~ px(345), Implication ~ px(365)) |>
  tab_style(style = cell_fill(color = "#EEEDFE"),
            locations = cells_body(rows = seq(1, 7, 2))) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13),
              column_labels.border.top.color = "#534AB7",
              column_labels.border.top.width = px(2))
```

---

## Reproducibility {#sec-repro}

```{r}
#| label: session-info
#| collapse: true
#| code-fold: true
#| code-summary: "Session info"

sessionInfo()
```

---

*Document rendered with [Quarto](https://quarto.org) on `r Sys.Date()`. Data: USITC Tariff Database via [TidyTuesday 2026 Week 17](https://github.com/rfordatascience/tidytuesday).*
