---
title: "US Agricultural Tariffs: A Comprehensive Analysis"
subtitle: "USITC Tariff Database · HTS Chapters 1–24 · TidyTuesday 2026 Week 17"
author:
  - name: "Erick K Yegon, PhD"
    email: "keyegon@gmail.com"
    orcid: "https://orcid.org/0000-0002-7055-4848"
date: today
date-format: "MMMM D, YYYY"

abstract: |
  This report is a **#TidyTuesday** submission for [Week 17, 2026](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-04-28/readme.md),
  exploring the USITC Tariff Database for agricultural products (HTS Chapters 1–24, 1997–2025).
  Using all four provided datasets — `tariff_agricultural`, `agreements`, `tariff_codes`,
  and `quantity_codes` — the analysis examines MFN rate trajectories, preferential savings
  under NAFTA/USMCA and bilateral agreements, tariff-line volatility, rate-type composition,
  and agreement coverage. Eleven publication-quality visualisations, two crosstalk-linked
  interactive widgets, seven statistical tables, and formal nonparametric tests support
  seven key descriptive findings. **No causal claims are made.**

keywords:
  - TidyTuesday
  - tariffs
  - trade policy
  - USITC
  - agricultural trade
  - NAFTA
  - USMCA
  - HTS
  - R
  - data-visualisation

# ── Format targets ────────────────────────────────────────────
format:
  html:
    theme:
      light: [flatly, custom.scss]
      dark:  [darkly,  custom.scss]
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-link: false    # disabled: code-link injects <a href> tags incompatible with gridtext/element_markdown
    code-summary: "Show / hide code"
    code-overflow: scroll
    fig-cap-location: bottom
    tbl-cap-location: top
    reference-location: margin
    footnotes-hover: true
    embed-resources: true
    smooth-scroll: true
    highlight-style: kate
    html-math-method: katex
    include-in-header:
      text: |
        <style>
        #reading-progress {
          position: fixed; top: 0; left: 0; height: 3px;
          background: #378ADD; z-index: 9999; width: 0%;
          transition: width 0.08s linear;
        }
        h1 { border-bottom: 2px solid #E6F1FB; padding-bottom: 6px; }
        h2 { border-left: 4px solid #378ADD; padding-left: 10px; }
        .metric-row { display: flex; gap: 16px; margin: 20px 0; flex-wrap: wrap; }
        .metric-card {
          flex: 1; min-width: 140px; background: #F1EFE8;
          border-radius: 8px; padding: 14px 18px; text-align: center;
        }
        .metric-card .val { font-size: 26px; font-weight: 700;
                            color: #185FA5; display: block; }
        .metric-card .lbl { font-size: 11px; color: #5F5E5A;
                            text-transform: uppercase; letter-spacing: .04em; }
        .tt-badge {
          display: inline-block; background: #534AB7; color: white;
          padding: 3px 10px; border-radius: 12px; font-size: 12px;
          font-weight: 600; font-family: monospace; margin-bottom: 8px;
        }
        .share-box {
          background: #F1EFE8; border: 1px solid #D3D1C7;
          border-radius: 8px; padding: 16px 20px; margin: 20px 0;
          font-size: 13px; line-height: 1.7;
        }
        .share-box a { color: #185FA5; }
        @media print {
          #reading-progress, .sidebar { display: none !important; }
          body { font-size: 12px; }
        }
        </style>
        <div id="reading-progress"></div>
        <script>
        document.addEventListener("scroll", function() {
          var s = document.documentElement;
          var pct = (s.scrollTop / (s.scrollHeight - s.clientHeight)) * 100;
          document.getElementById("reading-progress").style.width = pct + "%";
        });
        </script>
  pdf:
    documentclass: article
    papersize: letter
    geometry: [top=1in, bottom=1in, left=1in, right=1in]
    toc: true
    number-sections: true
    colorlinks: true
    linkcolor: "#185FA5"
    fig-pos: "H"
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{float}
        \usepackage{hyperref}
  revealjs:
    theme: moon
    slide-number: c/t
    chalkboard: true
    preview-links: true
    transition: fade
    progress: true
    footer: "#TidyTuesday · US Agricultural Tariffs · Week 17 2026"
    controls: true
    auto-animate: true

# ── Execution ─────────────────────────────────────────────────
execute:
  cache: true
  freeze: auto
  warning: false
  message: false
  echo: true

# ── Cross-references ──────────────────────────────────────────
crossref:
  fig-title: "Figure"
  tbl-title: "Table"
  fig-prefix: "Fig."
  tbl-prefix: "Tbl."
  sec-prefix: "§"
---

```{r}
#| label: setup
#| include: false
#| cache: false

knitr::opts_chunk$set(
  fig.align = "center",
  out.width = "100%",
  dpi       = 180,
  comment   = "#>"
)
```

```{r}
#| label: packages
#| cache: false

pkgs <- c(
  "tidyverse", "lubridate", "scales", "ggtext", "ggrepel",
  "patchwork", "gt", "gtExtras", "glue", "ggridges",
  "RColorBrewer", "forcats", "stringr", "janitor", "knitr",
  "kableExtra", "DT", "plotly", "crosstalk",
  "broom", "Kendall", "dunn.test"
)
installed  <- rownames(installed.packages())
to_install <- pkgs[!pkgs %in% installed]
if (length(to_install)) install.packages(to_install, quiet = TRUE)

suppressPackageStartupMessages({
  library(tidyverse); library(lubridate); library(scales)
  library(ggtext);    library(ggrepel);   library(patchwork)
  library(gt);        library(gtExtras);  library(glue)
  library(ggridges);  library(RColorBrewer); library(forcats)
  library(stringr);   library(janitor);   library(knitr)
  library(kableExtra);library(DT);        library(plotly)
  library(crosstalk); library(broom);     library(Kendall)
  library(dunn.test)
})
```

```{r}
#| label: theme-palette
#| cache: false

theme_tariff <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      text              = element_text(colour = "#2C2C2A"),
      plot.title        = element_markdown(size = rel(1.35), face = "bold",
                                           margin = margin(b = 6)),
      plot.subtitle     = element_markdown(size = rel(0.95), colour = "#5F5E5A",
                                           lineheight = 1.4, margin = margin(b = 12)),
      plot.caption      = element_markdown(size = rel(0.72), colour = "#888780",
                                           hjust = 0, margin = margin(t = 10)),
      plot.background   = element_rect(fill = "#FAFAF8", colour = NA),
      panel.background  = element_rect(fill = "#FAFAF8", colour = NA),
      panel.grid.major  = element_line(colour = "#D3D1C7", linewidth = 0.35),
      panel.grid.minor  = element_blank(),
      axis.title        = element_text(size = rel(0.85), colour = "#5F5E5A"),
      axis.text         = element_text(size = rel(0.80), colour = "#5F5E5A"),
      axis.ticks        = element_blank(),
      strip.text        = element_markdown(size = rel(0.85), face = "bold",
                                           colour = "#3C3489"),
      strip.background  = element_rect(fill = "#EEEDFE", colour = NA),
      legend.position   = "bottom",
      legend.title      = element_text(size = rel(0.82), colour = "#5F5E5A"),
      legend.text       = element_text(size = rel(0.78)),
      plot.margin       = margin(14, 14, 10, 14)
    )
}
theme_set(theme_tariff())

# Colour-blind-safe palette (Wong 2011)
pal_section <- c(
  "Live Animals & Animal Products" = "#D55E00",
  "Vegetable Products"             = "#009E73",
  "Animal/Vegetable Fats & Oils"   = "#CC79A7",
  "Prepared Foodstuffs"            = "#0072B2"
)

pal_agree <- c(
  "MFN"         = "#534AB7",
  "NAFTA/USMCA" = "#1D9E75",
  "Bilateral"   = "#D85A30",
  "Other GSP"   = "#BA7517"
)
```

---

## Introduction {#sec-intro}

<span class="tt-badge">#TidyTuesday · Week 17 · 2026-04-28</span>

::: {.callout-note icon="true"}
## About this dataset
This is a **[#TidyTuesday](https://github.com/rfordatascience/tidytuesday)** submission exploring US agricultural import tariffs from the **[USITC Tariff Database](https://www.usitc.gov/tariff_affairs/)**, covering all HTS Chapters 1–24 (live animals through beverages) with annual records from 1997 to 2025.

Four datasets are provided and **all four are used** in this analysis:

| Dataset | Role in this analysis |
|---|---|
| `tariff_agricultural` | Main data: one row per HTS-8 × agreement × date period |
| `agreements` | Lookup: agreement codes → full names, notes, eligibility |
| `tariff_codes` | Lookup: HTS-8 → product descriptions, quantity units, WTO binding |
| `quantity_codes` | Reference: unit codes (KG, NO, M2) → descriptions |

> ⚠️ **Important:** All patterns described below are **descriptive and correlational only**. There are many moderating variables — trade volumes, commodity market cycles, WTO binding commitments, exchange rates — not captured in this dataset. **No causal claims are made.**
:::

**Research questions (exploratory / descriptive):**

1. How have MFN tariff rates trended across the four agricultural HTS sections?
2. What rate differences exist between NAFTA/USMCA and MFN agreements?
3. Which specific tariff lines show the most rate variability over time?
4. Did the NAFTA → USMCA transition (2020) coincide with visible rate changes?
5. Is there an association between MFN rate levels and preferential agreement coverage?
6. Do rate distributions differ significantly across sections (nonparametric tests)?
7. Is there a detectable monotonic trend in MFN rates over time?

<div class="share-box">
📣 **Share this work:** Post with **#TidyTuesday** · Code on [GitHub](https://github.com/rfordatascience/tidytuesday/blob/main/sharing.md) · Dataset: [2026-04-28](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-04-28/readme.md)
</div>

---

## Data Ingestion & Validation {#sec-data}

```{r}
#| label: load-data
#| cache: true

base_url <- paste0(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/",
  "main/data/2026/2026-04-28/"
)

safe_read <- function(file) {
  readr::read_csv(paste0(base_url, file),
                  show_col_types = FALSE, progress = FALSE)
}

agreements          <- safe_read("agreements.csv")
quantity_codes      <- safe_read("quantity_codes.csv")
tariff_agricultural <- safe_read("tariff_agricultural.csv")
tariff_codes        <- safe_read("tariff_codes.csv")
```

```{r}
#| label: tbl-datasets
#| tbl-cap: "All four TidyTuesday datasets — dimensions after ingestion"

tibble(
  Dataset      = c("tariff_agricultural","agreements","quantity_codes","tariff_codes"),
  Rows         = c(nrow(tariff_agricultural), nrow(agreements),
                   nrow(quantity_codes),      nrow(tariff_codes)),
  Columns      = c(ncol(tariff_agricultural), ncol(agreements),
                   ncol(quantity_codes),      ncol(tariff_codes)),
  `Key join`   = c("hts8 + agreement + dates", "agreement",
                   "code", "hts8")
) |>
  gt() |>
  fmt_integer(columns = c(Rows, Columns)) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

```{r}
#| label: metric-cards
#| echo: false
#| results: asis

n_hts       <- n_distinct(tariff_agricultural$hts8)
n_agree_raw <- n_distinct(tariff_agricultural$agreement)
yr_min      <- year(min(as.Date(tariff_agricultural$begin_effective_date), na.rm = TRUE))
yr_max      <- year(max(as.Date(tariff_agricultural$end_effective_date),   na.rm = TRUE))
n_rows      <- scales::comma(nrow(tariff_agricultural))
n_qty_codes <- nrow(quantity_codes)

html_out <- paste0(
  '<div class="metric-row">',
  '<div class="metric-card"><span class="val">', n_rows,    '</span><span class="lbl">Tariff records</span></div>',
  '<div class="metric-card"><span class="val">', scales::comma(n_hts), '</span><span class="lbl">HTS-8 codes</span></div>',
  '<div class="metric-card"><span class="val">', n_agree_raw, '</span><span class="lbl">Agreement codes</span></div>',
  '<div class="metric-card"><span class="val">', yr_min, '–', yr_max, '</span><span class="lbl">Year span</span></div>',
  '<div class="metric-card"><span class="val">', n_qty_codes, '</span><span class="lbl">Quantity units</span></div>',
  '</div>'
)

knitr::asis_output(html_out)
```

::: {.column-margin}
**Quick facts**

HTS-8 codes: `r n_distinct(tariff_agricultural$hts8)`

Agreement codes: `r n_distinct(tariff_agricultural$agreement)`

Years: `r yr_min`–`r yr_max`

`ad_val_rate`: **decimal** (0.05 = 5%)
:::

### Data dictionary {#sec-dict}

::: {.callout-warning}
## `ad_val_rate` is stored as a decimal
Per the official data dictionary: `ad_val_rate` of `0.05` means **5%**, not 0.05%.
This is multiplied by 100 in all cleaning steps. `specific_rate` is in **US dollars per unit of quantity** (e.g., `0.05` = $0.05/kg).
:::

```{r}
#| label: tbl-data-dict
#| tbl-cap: "Key variables in tariff_agricultural.csv (official data dictionary)"

tribble(
  ~Variable,              ~Type,       ~Description,
  "hts8",                 "character", "8-digit HTS code — join with tariff_codes for descriptions",
  "begin_effective_date", "date",      "Start date for this rate period (rates can change mid-year)",
  "end_effective_date",   "date",      "End date — far-future dates (e.g. 2050-12-31) = no scheduled change",
  "agreement",            "character", "Agreement code — join with agreements for full name and notes",
  "rate_type_code",       "character", "Duty method (0=free, 7=ad valorem; see USITC td-codes.pdf)",
  "ad_val_rate",          "double",    "Ad valorem rate as DECIMAL — 0.05 = 5%",
  "specific_rate",        "double",    "Per-unit rate in USD — 0.05 = $0.05 per unit of quantity"
) |>
  gt() |>
  tab_header(title = md("**`tariff_agricultural` data dictionary**")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#FAEEDA"),
            locations = cells_body(rows = Variable == "ad_val_rate")) |>
  tab_style(style = cell_fill(color = "#FAEEDA"),
            locations = cells_body(rows = Variable == "specific_rate")) |>
  tab_options(table.font.size = px(12))
```

### Agreements reference {#sec-agreements}

```{r}
#| label: tbl-agreements
#| tbl-cap: "Full agreements lookup table — used to classify agree_group in cleaning"

agreements |>
  arrange(agreement) |>
  gt() |>
  tab_header(
    title    = md("**`agreements.csv` — all trade agreement codes**"),
    subtitle = md("*Asterisk/plus suffixes = restrictions or enhanced benefits*")
  ) |>
  cols_label(
    agreement       = "Code",
    agreement_full  = "Full name",
    agreement_notes = "Notes"
  ) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#EEEDFE"),
            locations = cells_body(rows = str_detect(agreement, "mfn"))) |>
  tab_style(style = cell_fill(color = "#EAF3DE"),
            locations = cells_body(rows = str_detect(agreement, "canada|mexico|usmca"))) |>
  sub_missing(missing_text = "—") |>
  tab_options(table.font.size = px(11))
```

### Quantity units in use {#sec-qty}

```{r}
#| label: tbl-qty
#| tbl-cap: "Quantity unit codes appearing in agricultural tariff lines (quantity_codes joined via tariff_codes)"

used_qty <- tariff_codes |>
  janitor::clean_names() |>
  select(quantity_1_code, quantity_2_code) |>
  pivot_longer(everything(), values_to = "code") |>
  filter(!is.na(code)) |>
  distinct(code)

quantity_codes |>
  semi_join(used_qty, by = "code") |>
  arrange(code) |>
  gt() |>
  tab_header(
    title    = md("**Quantity units from `quantity_codes.csv`**"),
    subtitle = md("*Filtered to codes used in HTS Ch. 1–24 tariff lines*")
  ) |>
  cols_label(code = "Unit code", description = "Description") |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(12))
```

---

## Data Cleaning & Feature Engineering {#sec-cleaning}

```{r}
#| label: cleaning

# ── 1. tariff_codes lookup: hts8 → description, qty units, WTO ──
tc_lookup <- tariff_codes |>
  janitor::clean_names() |>
  arrange(hts8, desc(earliest_effective_date)) |>
  distinct(hts8, .keep_all = TRUE) |>
  select(hts8, product_desc = description,
         qty1 = quantity_1_code, qty2 = quantity_2_code,
         wto  = wto_binding_code)

# ── 2. agreements classification (uses agreement column directly) ─
agree_classified <- agreements |>
  janitor::clean_names() |>
  mutate(
    agree_group = case_when(
      str_detect(agreement, "^mfn$")                                     ~ "MFN",
      str_detect(agreement, "canada|mexico|usmca")                       ~ "NAFTA/USMCA",
      str_detect(agreement,
                 "australia|chile|korea|colombia|peru|singapore|
                  bahrain|morocco|oman|panama|jordan|japan|israel_fta")   ~ "Bilateral",
      agreement == "col2"                                                ~ "Column 2",
      TRUE                                                               ~ "Other GSP"
    ),
    agree_group = factor(agree_group,
                         levels = c("MFN","NAFTA/USMCA","Bilateral","Other GSP","Column 2"))
  )

# ── 3. Clean tariff_agricultural ────────────────────────────────
# CRITICAL: ad_val_rate is decimal → multiply by 100 for percentage
tariff_clean <- tariff_agricultural |>
  janitor::clean_names() |>
  mutate(
    begin_date  = as.Date(begin_effective_date),
    end_date    = as.Date(end_effective_date),
    begin_year  = year(begin_date),
    end_year    = year(end_date),
    ad_val_pct  = ad_val_rate * 100,            # 0.05 → 5%
    hts_chapter = as.integer(str_sub(as.character(hts8), 1, 2)),
    section = case_when(
      hts_chapter <= 5  ~ "Live Animals & Animal Products",
      hts_chapter <= 14 ~ "Vegetable Products",
      hts_chapter <= 15 ~ "Animal/Vegetable Fats & Oils",
      hts_chapter <= 24 ~ "Prepared Foodstuffs",
      TRUE              ~ "Other"
    ),
    section = factor(section, levels = names(pal_section))
  ) |>
  left_join(tc_lookup,        by = "hts8") |>
  left_join(agree_classified, by = "agreement")

# ── 4. Annual panel (HTS8 × agreement × year 2000–2024) ─────────
YEARS <- 2000:2024

annual_panel <- tariff_clean |>
  distinct(hts8, agreement) |>
  crossing(year = YEARS) |>
  left_join(
    tariff_clean |>
      select(hts8, agreement, begin_year, end_year,
             ad_val_pct, specific_rate, hts_chapter, section,
             agree_group, agreement_full, agreement_notes,
             product_desc, qty1, qty2, wto),
    join_by(hts8, agreement, year >= begin_year, year <= end_year)
  ) |>
  filter(!is.na(ad_val_pct), ad_val_pct >= 0, ad_val_pct < 500) |>
  distinct(hts8, agreement, year, .keep_all = TRUE) |>
  filter(agree_group != "Column 2")   # exclude non-market economy rates

# ── 5. MFN-only subset ───────────────────────────────────────────
mfn_panel <- annual_panel |>
  filter(agree_group == "MFN") |>
  select(hts8, year, section, hts_chapter, mfn_rate = ad_val_pct,
         product_desc, qty1, wto)

cat("Annual panel rows:", scales::comma(nrow(annual_panel)), "\n")
cat("MFN panel rows   :", scales::comma(nrow(mfn_panel)), "\n")
cat("Distinct products:", scales::comma(n_distinct(mfn_panel$hts8)), "\n")
cat("MFN rate range   :", round(min(mfn_panel$mfn_rate), 2), "% –",
    round(max(mfn_panel$mfn_rate), 2), "%\n")
```

::: {.callout-tip}
## Pipeline design
- **Inequality join** (`join_by(year >= begin_year, year <= end_year)`) assigns the rate legally in force for each HTS × agreement × year — no rate-smearing across policy change dates.
- All four datasets joined: `tariff_agricultural` (base) + `agreements` (classification) + `tariff_codes` (descriptions) + `quantity_codes` (unit reference above).
- `ad_val_rate × 100` is the only non-obvious transformation required by the data dictionary.
:::

---

## Exploratory Data Analysis {#sec-eda}

::: {.callout-important appearance="minimal"}
**Reminder:** All patterns below are **descriptive only.** Observed associations do not imply causation. Many variables not in this dataset affect tariff levels and trade flows.
:::

### Section-level statistics {#sec-desc-stats}

```{r}
#| label: tbl-section-stats
#| tbl-cap: "MFN rate summary statistics by HTS section (2000–2024). WTO bound status from tariff_codes."

mfn_panel |>
  group_by(section) |>
  summarise(
    `Tariff lines` = n_distinct(hts8),
    `Obs. (n)`     = n(),
    `Mean rate`    = mean(mfn_rate, na.rm = TRUE),
    `Median`       = median(mfn_rate, na.rm = TRUE),
    `SD`           = sd(mfn_rate, na.rm = TRUE),
    `Min`          = min(mfn_rate, na.rm = TRUE),
    `Max`          = max(mfn_rate, na.rm = TRUE),
    `% zero-rated` = mean(mfn_rate == 0, na.rm = TRUE) * 100,
    `% WTO-bound`  = mean(wto == "B", na.rm = TRUE) * 100,
    .groups = "drop"
  ) |>
  gt() |>
  tab_header(
    title    = md("**MFN Rate Statistics by HTS Section**"),
    subtitle = md("*Ad valorem %. WTO bound (B) status from `tariff_codes.wto_binding_code`.*")
  ) |>
  cols_label(section = "Section") |>
  fmt_number(columns = c(`Mean rate`, Median, SD, Min, Max),
             decimals = 2, suffix = "%") |>
  fmt_number(columns = c(`% zero-rated`, `% WTO-bound`),
             decimals = 1, suffix = "%") |>
  fmt_integer(columns = c(`Tariff lines`, `Obs. (n)`)) |>
  data_color(columns = `Mean rate`,    palette = c("#E6F1FB","#185FA5")) |>
  data_color(columns = `% zero-rated`, palette = c("#185FA5","#E6F1FB")) |>
  data_color(columns = `% WTO-bound`,  palette = c("#E6F1FB","#009E73")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13),
              column_labels.border.top.color = "#534AB7",
              column_labels.border.top.width = px(2))
```

### Agreement group coverage {#sec-agree-cov}

```{r}
#| label: tbl-agreement-coverage
#| tbl-cap: "Coverage and average rates by agreement group — using agreement_notes to flag expired programs"

annual_panel |>
  group_by(agree_group) |>
  summarise(
    `Distinct products` = n_distinct(hts8),
    `Product-years`     = n(),
    `Mean rate`         = mean(ad_val_pct, na.rm = TRUE),
    `Median rate`       = median(ad_val_pct, na.rm = TRUE),
    `Expired/restricted` = sum(!is.na(agreement_notes) &
                               str_detect(agreement_notes,
                                          "expired|excluded"), na.rm = TRUE),
    .groups = "drop"
  ) |>
  gt() |>
  cols_label(agree_group = "Agreement group") |>
  fmt_number(columns = c(`Mean rate`, `Median rate`), decimals = 2, suffix = "%") |>
  fmt_integer(columns = c(`Distinct products`, `Product-years`, `Expired/restricted`)) |>
  data_color(columns = `Mean rate`, palette = c("#E6F1FB","#185FA5")) |>
  tab_footnote(
    footnote  = "Rows where agreement_notes contains 'expired' or 'excluded' (from agreements.csv)",
    locations = cells_column_labels(columns = `Expired/restricted`)
  ) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### Expired programs: rate distribution comparison {#sec-expired}

```{r}
#| label: fig-expired
#| fig-cap: "Rate distributions under active vs expired/restricted agreements. Uses agreement_notes from agreements.csv."
#| fig-width: 10
#| fig-height: 5

annual_panel |>
  mutate(
    status = case_when(
      str_detect(coalesce(agreement_notes,""), "expired")  ~ "Expired program",
      str_detect(coalesce(agreement_notes,""), "excluded") ~ "Restricted",
      agree_group == "MFN"                                 ~ "MFN (reference)",
      TRUE                                                 ~ "Active preferential"
    ),
    status = factor(status, levels = c("MFN (reference)","Active preferential",
                                       "Restricted","Expired program"))
  ) |>
  filter(ad_val_pct > 0, ad_val_pct < 100) |>
  ggplot(aes(x = ad_val_pct, fill = status, colour = status)) +
  geom_density(alpha = 0.35, linewidth = 0.7) +
  scale_fill_manual(
    values = c("MFN (reference)"     = "#534AB7",
               "Active preferential" = "#009E73",
               "Restricted"          = "#BA7517",
               "Expired program"     = "#D55E00"), name = NULL) +
  scale_colour_manual(
    values = c("MFN (reference)"     = "#534AB7",
               "Active preferential" = "#009E73",
               "Restricted"          = "#BA7517",
               "Expired program"     = "#D55E00"), name = NULL) +
  scale_x_continuous(labels = label_number(suffix = "%"), limits = c(0, 60)) +
  labs(
    title    = "Rate distributions by agreement status (uses agreement_notes field)",
    subtitle = "Expired = programs like ATPA (expired 2013). Restricted = CBI*, ATPA*.",
    x = "Ad valorem rate", y = "Density",
    caption  = "Source: USITC · agreements.agreement_notes · TidyTuesday 2026 W17 · #TidyTuesday"
  )
```

### OLS: MFN rate vs NAFTA/USMCA saving {#sec-regression}

```{r}
#| label: fig-regression
#| fig-cap: "Descriptive OLS: chapters with higher MFN rates tend to have larger NAFTA/USMCA differences. Not causal."
#| fig-width: 10
#| fig-height: 6

reg_data <- annual_panel |>
  filter(agree_group %in% c("MFN","NAFTA/USMCA")) |>
  group_by(hts_chapter, section, agree_group) |>
  summarise(avg = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = agree_group, values_from = avg) |>
  janitor::clean_names() |>
  filter(!is.na(mfn), !is.na(nafta_usmca)) |>
  mutate(saving = mfn - nafta_usmca)

model <- lm(saving ~ mfn, data = reg_data)
r2    <- glance(model)$r.squared
slope <- tidy(model)$estimate[2]

ggplot(reg_data, aes(x = mfn, y = saving, colour = section)) +
  geom_smooth(method = "lm", se = TRUE, colour = "#534AB7",
              fill = "#EEEDFE", linewidth = 0.8, alpha = 0.25) +
  geom_point(size = 4, alpha = 0.85) +
  geom_text_repel(aes(label = glue("Ch.{hts_chapter}")),
                  size = 2.8, max.overlaps = 20,
                  segment.colour = "#B4B2A9") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%")) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  annotate("richtext", x = -Inf, y = Inf, hjust = -0.05, vjust = 1.3,
           label = glue("R² = {round(r2,3)} · slope = {round(slope,2)} (descriptive)"),
           fill = NA, label.colour = NA, size = 3.2, colour = "#534AB7") +
  labs(
    title    = "Higher MFN rates are associated with larger NAFTA/USMCA rate differences",
    subtitle = "OLS across HTS chapters. 95% CI shown. **Descriptive only — not causal.**",
    x = "Mean MFN rate (%)", y = "Mean NAFTA/USMCA difference (pp)",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  )
```

```{r}
#| label: tbl-regression
#| tbl-cap: "OLS regression output (descriptive)"

bind_rows(
  tidy(model)   |> mutate(type = "Coefficients"),
  glance(model) |>
    select(r.squared, adj.r.squared, p.value, nobs) |>
    pivot_longer(everything(), names_to = "term", values_to = "estimate") |>
    mutate(type = "Model fit", std.error = NA_real_,
           statistic = NA_real_, p.value = NA_real_)
) |>
  gt() |>
  tab_header(title    = md("**OLS: NAFTA/USMCA saving ~ MFN rate (descriptive)**"),
             subtitle = md("*Unit: HTS chapter. No causal interpretation.*")) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  tab_row_group(label = "Coefficients", rows = type == "Coefficients") |>
  tab_row_group(label = "Model fit",    rows = type == "Model fit") |>
  cols_hide(type) |>
  tab_options(table.font.size = px(13))
```

### Lorenz / Gini: tariff rate inequality {#sec-gini}

```{r}
#| label: fig-gini
#| fig-cap: "Lorenz curves by section. Convexity = concentration of high rates in few tariff lines."
#| fig-width: 9
#| fig-height: 6

gini_fn <- function(x) {
  x <- sort(x[x > 0 & !is.na(x)])
  n <- length(x)
  if (n < 2) return(NA_real_)
  (2 * sum(seq_len(n) * x) / (n * sum(x))) - (n + 1) / n
}

lorenz_df <- mfn_panel |>
  filter(year == max(year)) |>
  group_by(section) |>
  filter(mfn_rate > 0) |>
  arrange(mfn_rate) |>
  mutate(
    cum_pop  = row_number() / n(),
    cum_rate = cumsum(mfn_rate) / sum(mfn_rate)
  ) |> ungroup()

gini_vals <- mfn_panel |>
  filter(year == max(year)) |>
  group_by(section) |>
  summarise(gini = gini_fn(mfn_rate), .groups = "drop") |>
  mutate(label = sprintf("Gini = %.3f", gini),
         x = 0.18, y = 0.80 - (as.integer(section) - 1) * 0.09)

ggplot(lorenz_df, aes(x = cum_pop, y = cum_rate, colour = section)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "#888780", linewidth = 0.6) +
  geom_line(linewidth = 1.2) +
  geom_label(data = gini_vals,
             aes(x = x, y = y,
                 label = paste0(str_wrap(as.character(section), 20), "\n", label),
                 colour = section),
             size = 2.5, fill = "#FAFAF8", label.size = 0.3,
             show.legend = FALSE) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_x_continuous(labels = label_percent()) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    title    = "Lorenz curves: concentration of MFN tariff rates by section",
    subtitle = glue("Latest year: {max(mfn_panel$year)}. Dashed = perfect equality."),
    x = "Cumulative share of tariff lines",
    y = "Cumulative share of total rate",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  )
```

---

## Visualisations {#sec-viz}

### MFN rate trajectories {#sec-fig-traj}

```{r}
#| label: fig-trajectories
#| fig-cap: "Mean MFN ad valorem rate by HTS section (2000–2024). Ribbon = ±1 SD. Grey = all sections."
#| fig-width: 11
#| fig-height: 7

section_yr <- mfn_panel |>
  group_by(section, year) |>
  summarise(
    mean_rate  = mean(mfn_rate, na.rm = TRUE),
    sd_rate    = sd(mfn_rate, na.rm = TRUE),
    n_products = n_distinct(hts8),
    .groups = "drop"
  ) |>
  filter(n_products >= 5) |>
  mutate(lo = pmax(mean_rate - sd_rate, 0), hi = mean_rate + sd_rate)

bg_lines <- section_yr |> rename(bg_section = section)

ggplot() +
  geom_line(data = bg_lines,
            aes(x = year, y = mean_rate, group = bg_section),
            colour = "#D3D1C7", linewidth = 0.7, alpha = 0.6) +
  geom_ribbon(data = section_yr,
              aes(x = year, ymin = lo, ymax = hi, fill = section), alpha = 0.18) +
  geom_line(data = section_yr,
            aes(x = year, y = mean_rate, colour = section), linewidth = 1.3) +
  geom_point(data = section_yr,
             aes(x = year, y = mean_rate, colour = section),
             size = 1.6, shape = 21, fill = "white", stroke = 1.0) +
  scale_colour_manual(values = pal_section, guide = "none") +
  scale_fill_manual(values = pal_section, guide = "none") +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  scale_y_continuous(labels = label_number(suffix = "%"), limits = c(0, NA)) +
  facet_wrap(~section, ncol = 2, scales = "free_y") +
  labs(
    title    = "MFN tariff trends by agricultural section (2000–2024)",
    subtitle = "Mean rate ± 1 SD (ribbon). Grey = all sections for context.",
    x = NULL, y = "Average MFN rate",
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 W17 · #TidyTuesday"
  )
```

::: {.callout-note appearance="minimal"}
**Descriptive observation:** Live Animals & Animal Products show persistently higher average rates. Vegetable Products show a mild downward trend in the 2010s. These patterns warrant further investigation but should not be interpreted causally from this data alone.
:::

### Year-over-year change heatmap {#sec-fig-yoy}

```{r}
#| label: fig-yoy
#| fig-cap: "Year-over-year change in mean MFN rate per section. Red = increase, blue = decrease."
#| fig-height: 5.5

mfn_panel |>
  group_by(section, year) |>
  summarise(mean_rate = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  arrange(section, year) |>
  group_by(section) |>
  mutate(yoy_change = mean_rate - lag(mean_rate)) |>
  ungroup() |>
  filter(!is.na(yoy_change)) |>
  ggplot(aes(x = year, y = fct_rev(section), fill = yoy_change)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = ifelse(abs(yoy_change) >= 0.3,
                               sprintf("%+.1f", yoy_change), "")),
            size = 2.5, colour = "white", fontface = "bold") +
  scale_fill_gradient2(
    low = "#185FA5", mid = "#F1EFE8", high = "#D85A30", midpoint = 0,
    name = "YoY Δ (pp)",
    guide = guide_colourbar(barwidth = 10, barheight = 0.5,
                            title.position = "top", title.hjust = 0.5)
  ) +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  labs(
    title    = "Year-over-year MFN rate changes by section",
    subtitle = "Red = increase · Blue = decrease · Labels for |Δ| ≥ 0.3 pp.",
    x = NULL, y = NULL,
    caption  = "Source: USITC Tariff Database · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(panel.grid = element_blank(), axis.text.y = element_markdown(size = 9))
```

### Agreement coverage heatmap {#sec-fig-heatmap}

```{r}
#| label: fig-heatmap
#| fig-cap: "Average ad valorem rate by HTS chapter and agreement group. Darker = higher rates."
#| fig-width: 10
#| fig-height: 13

annual_panel |>
  mutate(chapter_label = glue("Ch. {str_pad(hts_chapter, 2, 'left', '0')}")) |>
  group_by(section, hts_chapter, chapter_label, agree_group) |>
  summarise(avg_rate = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  filter(agree_group %in% c("MFN","NAFTA/USMCA","Bilateral")) |>
  mutate(
    agree_group   = factor(agree_group, levels = c("MFN","NAFTA/USMCA","Bilateral")),
    chapter_label = fct_reorder(chapter_label, hts_chapter)
  ) |>
  ggplot(aes(x = agree_group, y = fct_rev(chapter_label), fill = avg_rate)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(avg_rate >= 1, sprintf("%.1f%%", avg_rate), "")),
            size = 2.5, colour = "white", fontface = "bold") +
  scale_fill_gradientn(
    colours  = c("#E6F1FB","#378ADD","#185FA5","#042C53"),
    na.value = "#F1EFE8", name = "Avg. rate",
    labels   = label_number(suffix = "%"),
    guide    = guide_colourbar(barwidth = 10, barheight = 0.5,
                               title.position = "top", title.hjust = 0.5)
  ) +
  facet_grid(section ~ ., scales = "free_y", space = "free_y") +
  labs(
    title    = "Tariff rates by HTS chapter and agreement group",
    subtitle = "Each cell = mean ad valorem rate for that chapter–agreement pair.",
    x = "Agreement group", y = "HTS chapter",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(panel.grid   = element_blank(),
        axis.text.y  = element_text(size = 7),
        strip.text.y = element_text(angle = 0, hjust = 0, size = 7))
```

### MFN vs NAFTA/USMCA dumbbell {#sec-fig-dumbbell}

```{r}
#| label: fig-dumbbell
#| fig-cap: "MFN (filled) vs NAFTA/USMCA (open) average rates per chapter. Segment = pp difference."
#| fig-width: 10
#| fig-height: 8

dumbbell_data <- annual_panel |>
  filter(agree_group %in% c("MFN","NAFTA/USMCA")) |>
  group_by(hts_chapter, section, agree_group) |>
  summarise(avg_rate = mean(ad_val_pct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = agree_group, values_from = avg_rate) |>
  clean_names() |>
  filter(!is.na(mfn), !is.na(nafta_usmca)) |>
  mutate(
    saving        = mfn - nafta_usmca,
    chapter_label = glue("Ch. {str_pad(hts_chapter, 2, 'left', '0')}"),
    chapter_label = fct_reorder(chapter_label, saving)
  ) |>
  filter(saving > 0)

ggplot(dumbbell_data) +
  geom_segment(aes(x = nafta_usmca, xend = mfn,
                   y = chapter_label, yend = chapter_label, colour = section),
               linewidth = 1.8, alpha = 0.40) +
  geom_point(aes(x = nafta_usmca, y = chapter_label, colour = section),
             size = 3.8, shape = 21, fill = "white", stroke = 1.2) +
  geom_point(aes(x = mfn, y = chapter_label, colour = section), size = 3.8) +
  geom_text(aes(x = mfn + 0.25, y = chapter_label,
                label = sprintf("−%.1f pp", saving)),
            hjust = 0, size = 2.9, colour = "#5F5E5A") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0.01, 0.16))) +
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(
    title    = "NAFTA/USMCA vs MFN: average rate differences by chapter",
    subtitle = "Open = NAFTA/USMCA · Filled = MFN · Label = pp difference (descriptive).",
    x = "Average ad valorem rate", y = NULL,
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(panel.grid.major.y = element_blank())
```

### Tariff volatility — top 30 lines {#sec-fig-vol}

```{r}
#| label: fig-volatility
#| fig-cap: "Top 30 MFN tariff lines by rate SD. Labels use product_desc from tariff_codes.csv."
#| fig-width: 12
#| fig-height: 10

mfn_panel |>
  group_by(hts8, section, product_desc) |>
  filter(n() >= 5) |>
  summarise(rate_sd  = sd(mfn_rate, na.rm = TRUE),
            mean_mfn = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  filter(mean_mfn > 0) |>
  slice_max(rate_sd, n = 30) |>
  mutate(
    label = coalesce(str_trunc(product_desc, 48), as.character(hts8)),
    label = fct_reorder(label, rate_sd)
  ) |>
  ggplot(aes(x = rate_sd, y = label, colour = section)) +
  geom_segment(aes(x = 0, xend = rate_sd, y = label, yend = label),
               linewidth = 0.9, alpha = 0.50) +
  geom_point(size = 3.6) +
  geom_text(aes(label = sprintf("%.1f%%", rate_sd)),
            hjust = -0.3, size = 2.7, colour = "#5F5E5A") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = "%"),
                     expand = expansion(mult = c(0, 0.18))) +
  labs(
    title    = "Top 30 most volatile MFN tariff lines",
    subtitle = "Labels from tariff_codes.description. Ranked by SD of annual MFN rate, 2000–2024.",
    x = "Rate SD (pp)", y = NULL,
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_text(size = 7.5))
```

### NAFTA → USMCA transition {#sec-fig-nafta}

```{r}
#| label: fig-nafta
#| fig-cap: "NAFTA/USMCA preferential rate trajectories. Dashed = USMCA effective date (2020)."
#| fig-width: 11
#| fig-height: 6

nafta_section <- annual_panel |>
  filter(agree_group == "NAFTA/USMCA") |>
  group_by(year, section) |>
  summarise(
    mean_rate = mean(ad_val_pct, na.rm = TRUE),
    lo        = quantile(ad_val_pct, 0.25, na.rm = TRUE),
    hi        = quantile(ad_val_pct, 0.75, na.rm = TRUE),
    .groups   = "drop"
  )

ggplot(nafta_section, aes(x = year, colour = section, fill = section)) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             colour = "#534AB7", linewidth = 0.7, alpha = 0.7) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.14, linewidth = 0) +
  geom_line(aes(y = mean_rate), linewidth = 1.2) +
  geom_point(aes(y = mean_rate), size = 1.7, shape = 21, fill = "white") +
  annotate("richtext", x = 2020.4, y = Inf, hjust = 0, vjust = 1.4,
           size = 3.2, label = "**USMCA** replaces NAFTA",
           fill = NA, label.colour = NA, colour = "#534AB7") +
  scale_colour_manual(values = pal_section, name = NULL) +
  scale_fill_manual(values = pal_section, name = NULL) +
  scale_x_continuous(breaks = seq(2000, 2024, 4)) +
  scale_y_continuous(labels = label_number(suffix = "%"), limits = c(0, NA)) +
  labs(
    title    = "NAFTA → USMCA: preferential rate trajectories",
    subtitle = "agreements.csv codes: 'canada'/'mexico' for NAFTA, 'usmca' for USMCA. Mean ± IQR.\nRates near 0% after ~2008 reflect completed NAFTA phase-out schedules.",
    x = NULL, y = "Average preferential rate (%)",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(legend.position = "top")
```

### Biggest rate movers {#sec-fig-movers}

```{r}
#| label: fig-movers
#| fig-cap: "Top 15 MFN rate changes. Labels from tariff_codes.description via product_desc."
#| fig-width: 13
#| fig-height: 10

big_movers <- mfn_panel |>
  group_by(hts8, section, product_desc) |>
  filter(n_distinct(year) >= 10) |>
  summarise(
    rate_start = mean(mfn_rate[year == min(year)], na.rm = TRUE),
    rate_end   = mean(mfn_rate[year == max(year)], na.rm = TRUE),
    change     = rate_end - rate_start, .groups = "drop"
  ) |>
  filter(!is.na(change)) |>
  mutate(direction = if_else(change > 0, "Increase", "Decrease")) |>
  group_by(direction) |>
  slice_max(abs(change), n = 15) |>
  ungroup() |>
  mutate(
    label = coalesce(str_trunc(product_desc, 42), as.character(hts8)),
    label = fct_reorder(label, change)
  )

ggplot(big_movers, aes(x = change, y = label, fill = section)) +
  geom_col(alpha = 0.82, width = 0.72) +
  geom_vline(xintercept = 0, colour = "#2C2C2A", linewidth = 0.65) +
  geom_text(aes(label = sprintf("%+.1f pp", change),
                hjust = if_else(change > 0, -0.15, 1.15)),
            size = 2.8, colour = "#2C2C2A") +
  scale_fill_manual(values = pal_section, name = "Section") +
  scale_x_continuous(labels = label_number(suffix = " pp", style_positive = "plus"),
                     expand = expansion(mult = 0.20)) +
  labs(
    title    = "Biggest MFN rate changes (earliest → latest observation)",
    subtitle = "Top 15 increases and decreases. Labels from tariff_codes.description.",
    x = "Change (pp)", y = NULL,
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y        = element_markdown(size = 7.2))
```

### Agreements vs MFN rate — bubble {#sec-fig-bubble}

```{r}
#| label: fig-bubble
#| fig-cap: "Association between MFN rate level and preferential agreement group coverage — descriptive only."
#| fig-width: 10
#| fig-height: 7

bubble_chapter <- annual_panel |>
  group_by(hts8, hts_chapter, section) |>
  summarise(
    n_agreements = n_distinct(agree_group[agree_group != "MFN"]),
    mean_mfn     = mean(ad_val_pct[agree_group == "MFN"], na.rm = TRUE),
    .groups      = "drop"
  ) |>
  filter(!is.na(mean_mfn), mean_mfn > 0) |>
  group_by(hts_chapter, section) |>
  summarise(
    n_agreements = mean(n_agreements, na.rm = TRUE),
    mean_mfn     = mean(mean_mfn, na.rm = TRUE),
    n_products   = n(), .groups = "drop"
  ) |>
  mutate(chapter_label = glue("Ch.{hts_chapter}"))

ggplot(bubble_chapter,
       aes(x = n_agreements, y = mean_mfn,
           size = n_products, colour = section)) +
  geom_point(alpha = 0.72) +
  geom_text_repel(aes(label = chapter_label), size = 2.7, max.overlaps = 20,
                  segment.colour = "#B4B2A9", segment.size    = 0.4) +
  scale_size_area(max_size = 14, name = "# products") +
  scale_colour_manual(values = pal_section, name = "Section") +
  scale_x_continuous(breaks = 0:5, minor_breaks = NULL) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  guides(size   = guide_legend(override.aes = list(colour = "#888780")),
         colour = guide_legend(override.aes = list(size = 4))) +
  labs(
    title    = "Association between MFN rate level and agreement coverage (descriptive)",
    subtitle = "Bubble = HTS chapter · Size = tariff lines · X = avg. agreement groups.",
    x = "Average number of preferential agreement groups",
    y = "Mean MFN tariff rate",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  )
```

### Tariff type composition {#sec-fig-comp}

```{r}
#| label: fig-composition
#| fig-cap: "Share of MFN lines by rate structure. Specific = $/unit from specific_rate column."
#| fig-width: 11
#| fig-height: 7

type_pal <- c("Ad valorem only" = "#378ADD", "Specific only" = "#D85A30",
              "Compound (both)" = "#BA7517", "Zero / free"   = "#D3D1C7")

annual_panel |>
  filter(agree_group == "MFN") |>
  mutate(
    rate_type = case_when(
      !is.na(specific_rate) & specific_rate > 0 &
        !is.na(ad_val_pct)   & ad_val_pct   > 0  ~ "Compound (both)",
      !is.na(specific_rate) & specific_rate > 0  ~ "Specific only",
      !is.na(ad_val_pct)   & ad_val_pct   > 0  ~ "Ad valorem only",
      TRUE                                       ~ "Zero / free"
    )
  ) |>
  count(year, section, rate_type) |>
  group_by(year, section) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = year, y = share, fill = rate_type)) +
  geom_area(alpha = 0.85, position = "stack") +
  scale_fill_manual(values = type_pal, name = "Tariff type") +
  scale_x_continuous(breaks = seq(2000, 2024, 6)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  facet_wrap(~section, ncol = 2) +
  labs(
    title    = "MFN tariff rate structure composition by section",
    subtitle = "Specific = per-unit USD (specific_rate). Compound = both ad valorem and specific.",
    x = NULL, y = "Share of tariff lines",
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(legend.position = "bottom")
```

### Rate distribution ridgelines {#sec-fig-ridges}

```{r}
#| label: fig-ridges
#| fig-cap: "MFN ad valorem rate distributions every four years. X-axis capped at 60%."
#| fig-width: 11
#| fig-height: 8

mfn_panel |>
  filter(year %in% seq(2000, 2024, 4), mfn_rate > 0, mfn_rate <= 100) |>
  mutate(year_f = factor(year)) |>
  ggplot(aes(x = mfn_rate, y = year_f, fill = section)) +
  geom_density_ridges(alpha = 0.65, scale = 1.8,
                      rel_min_height = 0.01, bandwidth = 1.5) +
  scale_fill_manual(values = pal_section, name = NULL) +
  scale_x_continuous(labels = label_number(suffix = "%"), limits = c(0, 60)) +
  facet_wrap(~section, ncol = 2) +
  labs(
    title    = "Distribution of MFN tariff rates over time (every 4 years)",
    subtitle = "Right-skew visible in all sections. X-axis capped at 60%.",
    x = "MFN ad valorem rate", y = NULL,
    caption  = "Source: USITC · TidyTuesday 2026 W17 · #TidyTuesday"
  ) +
  theme(legend.position = "none")
```

---

## Statistical Tests {#sec-stats}

::: {.callout-important appearance="minimal"}
Statistical significance does not imply practical importance or causation. Tests below assess consistency with random variation only.
:::

### Kruskal-Wallis {#sec-kruskal}

```{r}
#| label: tbl-kruskal
#| tbl-cap: "Kruskal-Wallis test: do MFN rate distributions differ across HTS sections?"

kw <- kruskal.test(mfn_rate ~ section, data = mfn_panel)

tibble(
  Test      = "Kruskal-Wallis rank-sum",
  Statistic = round(kw$statistic, 2),
  df        = kw$parameter,
  `p-value` = formatC(kw$p.value, format = "e", digits = 3),
  Conclusion = if_else(kw$p.value < 0.05,
                       "Rate distributions differ across sections (p < 0.05)",
                       "No significant difference detected")
) |>
  gt() |>
  tab_header(title    = md("**Kruskal-Wallis test**"),
             subtitle = md("*H₀: all sections share the same rate distribution*")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### Dunn post-hoc {#sec-dunn}

```{r}
#| label: tbl-dunn
#| tbl-cap: "Dunn pairwise comparisons (Bonferroni correction)"

dunn_res <- dunn.test(
  mfn_panel$mfn_rate, mfn_panel$section,
  method = "bonferroni", alpha = 0.05, kw = FALSE, label = TRUE
)

tibble(
  Comparison   = dunn_res$comparisons,
  Z            = round(dunn_res$Z, 3),
  `p-adjusted` = round(dunn_res$P.adjusted, 4),
  Significant  = if_else(dunn_res$P.adjusted < 0.05, "Yes ✓", "No")
) |>
  gt() |>
  tab_header(title    = md("**Dunn post-hoc (Bonferroni)**"),
             subtitle = md("*Pairwise section comparisons*")) |>
  tab_style(style = cell_fill(color = "#EAF3DE"),
            locations = cells_body(rows = Significant == "Yes ✓")) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

### Mann-Kendall trend test {#sec-mk}

```{r}
#| label: tbl-mk
#| tbl-cap: "Mann-Kendall monotonic trend test by section (2000–2024)"

mfn_panel |>
  group_by(section, year) |>
  summarise(mean_rate = mean(mfn_rate, na.rm = TRUE), .groups = "drop") |>
  group_by(section) |>
  summarise(
    tau     = MannKendall(mean_rate)$tau[[1]],
    p_value = MannKendall(mean_rate)$sl[[1]],
    trend   = case_when(
      p_value >= 0.05           ~ "No monotonic trend detected",
      tau < 0 & p_value < 0.05 ~ "Significant decrease ↓",
      tau > 0 & p_value < 0.05 ~ "Significant increase ↑"
    ),
    .groups = "drop"
  ) |>
  gt() |>
  tab_header(title    = md("**Mann-Kendall trend test (2000–2024)**"),
             subtitle = md("*H₀: no monotonic trend in annual mean MFN rate*")) |>
  cols_label(section = "Section", tau = "Kendall τ",
             p_value = "p-value", trend = "Conclusion") |>
  fmt_number(columns = c(tau, p_value), decimals = 4) |>
  tab_style(style = cell_fill(color = "#E6F1FB"),
            locations = cells_body(rows = str_detect(trend, "decrease"))) |>
  tab_style(style = cell_fill(color = "#FAEEDA"),
            locations = cells_body(rows = str_detect(trend, "increase"))) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13))
```

---

## Interactive Exploration {#sec-interactive}

### Linked chapter explorer {#sec-crosstalk}

```{r}
#| label: crosstalk-filter

shared_chapter <- mfn_panel |>
  group_by(hts_chapter, section) |>
  summarise(
    Products  = n_distinct(hts8),
    Mean_rate = round(mean(mfn_rate, na.rm = TRUE), 2),
    Median    = round(median(mfn_rate, na.rm = TRUE), 2),
    SD        = round(sd(mfn_rate, na.rm = TRUE), 2),
    Max       = round(max(mfn_rate, na.rm = TRUE), 2),
    .groups   = "drop"
  ) |>
  rename(Chapter = hts_chapter, Section = section) |>
  SharedData$new(key = ~Chapter)

filter_select("sec_filter", "Filter by section:", shared_chapter, ~Section)
```

```{r}
#| label: tbl-crosstalk
#| tbl-cap: "MFN rate statistics by HTS chapter — select rows to highlight in scatter below"

DT::datatable(
  shared_chapter, rownames = FALSE,
  extensions = "Buttons",
  options = list(dom = "Bfrtip", buttons = c("copy","csv"),
                 pageLength = 12, scrollX = TRUE)
) |>
  DT::formatStyle(
    "Mean_rate",
    background         = DT::styleColorBar(c(0, 30), "#378ADD"),
    backgroundSize     = "100% 90%",
    backgroundRepeat   = "no-repeat",
    backgroundPosition = "center"
  )
```

```{r}
#| label: fig-crosstalk
#| fig-cap: "Linked scatter — select table rows to highlight chapters."

plot_ly(shared_chapter,
        x = ~Mean_rate, y = ~SD, size = ~Products, color = ~Section,
        colors = unname(pal_section), type = "scatter", mode = "markers",
        marker = list(opacity = 0.75, sizemode = "diameter"),
        hovertemplate = "<b>Ch %{customdata}</b><br>Mean: %{x:.1f}%<br>SD: %{y:.1f}%<extra></extra>",
        customdata = ~Chapter
) |>
  layout(
    title  = list(text = "Chapter rate level vs volatility (linked to table)"),
    xaxis  = list(title = "Mean MFN rate (%)", gridcolor = "#D3D1C7"),
    yaxis  = list(title = "Rate SD (pp)",       gridcolor = "#D3D1C7"),
    legend = list(orientation = "h", y = -0.2),
    paper_bgcolor = "#FAFAF8", plot_bgcolor = "#FAFAF8"
  ) |>
  highlight(on = "plotly_selected", off = "plotly_deselect",
            color = "#D85A30", persistent = TRUE)
```

### Interactive trajectories with IQR {#sec-plotly}

```{r}
#| label: fig-plotly
#| fig-cap: "Interactive MFN trajectories with IQR bands — unified hover tooltip."

section_yr_plotly <- mfn_panel |>
  group_by(section, year) |>
  summarise(
    mean_rate = mean(mfn_rate, na.rm = TRUE),
    lo        = quantile(mfn_rate, 0.25, na.rm = TRUE),
    hi        = quantile(mfn_rate, 0.75, na.rm = TRUE),
    .groups   = "drop"
  )

p <- plot_ly()
for (sec in levels(section_yr_plotly$section)) {
  d   <- filter(section_yr_plotly, section == sec)
  col <- pal_section[sec]
  p   <- add_ribbons(p, data = d, x = ~year, ymin = ~lo, ymax = ~hi,
                     fillcolor = paste0(col, "22"), line = list(width = 0),
                     name = paste0(sec, " IQR"), legendgroup = sec,
                     showlegend = FALSE, hoverinfo = "skip")
  p   <- add_lines(p, data = d, x = ~year, y = ~mean_rate,
                   line = list(color = col, width = 2.5),
                   name = sec, legendgroup = sec,
                   hovertemplate = paste0("<b>", sec, "</b><br>Year: %{x}<br>",
                                          "Mean: %{y:.2f}%<extra></extra>"))
}
p |> layout(
  title     = list(text = "MFN rate trajectories with IQR bands"),
  xaxis     = list(title = "", tickvals = seq(2000, 2024, 4), gridcolor = "#D3D1C7"),
  yaxis     = list(title = "Average MFN rate (%)", rangemode = "tozero",
                   gridcolor = "#D3D1C7"),
  legend    = list(orientation = "h", y = -0.2),
  hovermode = "x unified",
  paper_bgcolor = "#FAFAF8", plot_bgcolor = "#FAFAF8"
)
```

---




## Composite Dashboard {#sec-dashboard}

```{r}
#| label: fig-dashboard
#| fig-cap: "Four-panel summary dashboard — #TidyTuesday 2026 Week 17 · © Erick K Yegon, PhD (@eyegonk)"
#| fig-width: 18
#| fig-height: 12
#| column: screen
#| dev: ragg_png
#| dpi: 300

# ╔══════════════════════════════════════════════════════════════════════════╗
#   DASHBOARD THEME & PALETTE
#   Dark-mode navy theme optimised for Twitter/social media sharing
#   Colour-blind-safe hues boosted for dark backgrounds
# ╚══════════════════════════════════════════════════════════════════════════╝

# ── Dark mini-theme (overrides the global theme_tariff for dashboard only) ──
theme_dash <- function() {
  theme_tariff() +
    theme(
      # Page & panel backgrounds
      plot.background   = element_rect(fill = "#0F1923", colour = NA),
      panel.background  = element_rect(fill = "#141E2B", colour = NA),
      # Subtle grid — visible but not distracting on dark bg
      panel.grid.major  = element_line(colour = "#1E2D3D", linewidth = 0.35),
      panel.grid.minor  = element_blank(),
      # Axis text — muted steel blue
      axis.text         = element_text(colour = "#7A8FA6", size = 7.5,
                                       family = "sans"),
      axis.title        = element_text(colour = "#7A8FA6", size = 8),
      axis.ticks        = element_blank(),
      # Facet strips
      strip.background  = element_rect(fill = "#1A2840", colour = NA),
      strip.text        = element_markdown(colour = "#C8D8E8", face = "bold",
                                           size = 7.5),
      # Per-panel title & subtitle
      plot.title        = element_markdown(colour = "#E8F4FD", face = "bold",
                                           size = 11, margin = margin(b = 3)),
      plot.subtitle     = element_markdown(colour = "#7A8FA6", size = 8,
                                           lineheight = 1.3,
                                           margin = margin(b = 8)),
      plot.margin       = margin(12, 14, 10, 14),
      # Legend
      legend.background = element_rect(fill = "#141E2B", colour = NA),
      legend.text       = element_text(colour = "#C8D8E8", size = 7.5),
      legend.title      = element_text(colour = "#7A8FA6", size = 7.5),
      legend.key        = element_rect(fill = NA, colour = NA),
      legend.position   = "none"
    )
}

# ── Dark-mode palette ──────────────────────────────────────────────────────
# Same hue family as pal_section but boosted saturation/brightness for navy bg
pal_dark <- c(
  "Live Animals & Animal Products" = "#FF8C5A",   # warm orange
  "Vegetable Products"             = "#3DDC84",   # vivid green
  "Animal/Vegetable Fats & Oils"   = "#E060E0",   # magenta
  "Prepared Foodstuffs"            = "#4DA8FF"    # sky blue
)

# ── Copyright string ────────────────────────────────────────────────────────
# IMPORTANT: element_markdown() uses gridtext which does NOT support <a> tags.
# Only plain text + Unicode is safe here. Auto-updates year on each render.
copyright_str <- glue(
  "\u00a9 {format(Sys.Date(), '%Y')} Erick K Yegon, PhD  \u00b7  ",
  "Data: USITC Tariff Database via TidyTuesday 2026 W17  \u00b7  ",
  "keyegon@gmail.com  \u00b7  @eyegonk  \u00b7  ",
  "#rstats #dataviz #TidyTuesday"
)

# ╔══════════════════════════════════════════════════════════════════════════╗
#   PANEL 1 — MFN RATE TRAJECTORIES
#   Small multiples + uncertainty ribbons + end-of-series inline labels
# ╚══════════════════════════════════════════════════════════════════════════╝

# Build end-of-series annotation data (2024 values per section)
end_labels <- section_yr |>
  group_by(section) |>
  filter(year == max(year)) |>
  ungroup() |>
  mutate(label = sprintf("%.1f%%", mean_rate))

p1 <- ggplot() +
  # Grey context lines — all four sections faintly in every facet
  geom_line(
    data    = bg_lines,
    mapping = aes(x = year, y = mean_rate, group = bg_section),
    colour  = "#1E2D3D", linewidth = 0.9
  ) +
  # +/-1 SD uncertainty ribbon
  geom_ribbon(
    data    = section_yr,
    mapping = aes(x = year, ymin = lo, ymax = hi, fill = section),
    alpha   = 0.22
  ) +
  # Mean trajectory line
  geom_line(
    data      = section_yr,
    mapping   = aes(x = year, y = mean_rate, colour = section),
    linewidth = 1.25
  ) +
  # Terminal dot at 2024
  geom_point(
    data    = end_labels,
    mapping = aes(x = year, y = mean_rate, colour = section),
    size = 2.8
  ) +
  # Inline rate label at series end — plain text, no markdown needed here
  geom_text(
    data    = end_labels,
    mapping = aes(x = year + 0.5, y = mean_rate,
                  colour = section, label = label),
    hjust = 0, size = 2.6, fontface = "bold"
  ) +
  scale_colour_manual(values = pal_dark, guide = "none") +
  scale_fill_manual(values   = pal_dark, guide = "none") +
  scale_x_continuous(
    breaks = seq(2000, 2024, 8),
    expand = expansion(mult = c(0.01, 0.18))   # extra room for end labels
  ) +
  scale_y_continuous(
    labels = label_number(suffix = "%"),
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  facet_wrap(~section, ncol = 2, scales = "free_y") +
  labs(
    title    = "\u2460 MFN RATE TRAJECTORIES",
    subtitle = "Mean \u00b1 1 SD ribbon \u00b7 2000\u20132024 \u00b7 Grey = all sections for context",
    x = NULL,
    y = "Mean MFN rate"
  ) +
  theme_dash()

# ╔══════════════════════════════════════════════════════════════════════════╗
#   PANEL 2 — DUMBBELL: MFN vs NAFTA/USMCA
#   Top 14 chapters by percentage-point saving — sorted ascending
# ╚══════════════════════════════════════════════════════════════════════════╝

db_plot <- dumbbell_data |>
  slice_max(saving, n = 14) |>
  mutate(
    chapter_label = fct_reorder(chapter_label, saving),
    saving_label  = sprintf("\u2212%.1f pp", saving)   # proper Unicode minus
  )

p2 <- ggplot(db_plot) +
  # Connecting segment — thickness encodes magnitude of saving
  geom_segment(
    aes(x = nafta_usmca, xend = mfn,
        y = chapter_label, yend = chapter_label,
        colour = section),
    linewidth = 2.2, alpha = 0.35
  ) +
  # Open circle = NAFTA/USMCA preferential rate
  geom_point(
    aes(x = nafta_usmca, y = chapter_label, colour = section),
    size = 3.4, shape = 21, fill = "#141E2B", stroke = 1.5
  ) +
  # Filled circle = MFN rate
  geom_point(
    aes(x = mfn, y = chapter_label, colour = section),
    size = 3.4
  ) +
  # Percentage-point saving label to the right of MFN dot
  geom_text(
    aes(x = mfn + 0.35, y = chapter_label, label = saving_label),
    hjust = 0, size = 2.6, colour = "#7A8FA6", fontface = "bold"
  ) +
  scale_colour_manual(values = pal_dark, guide = "none") +
  scale_x_continuous(
    labels = label_number(suffix = "%"),
    expand = expansion(mult = c(0.01, 0.22))
  ) +
  labs(
    title    = "\u2461 MFN vs NAFTA/USMCA RATES",
    subtitle = "\u25cf Filled = MFN rate  \u00b7  \u25cb Open = Preferential  \u00b7  Label = pp saving",
    x = "Average ad valorem rate",
    y = NULL
  ) +
  theme_dash() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y        = element_text(size = 6.8, colour = "#C8D8E8")
  )

# ╔══════════════════════════════════════════════════════════════════════════╗
#   PANEL 3 — NAFTA → USMCA TRANSITION
#   Ribbon + shaded USMCA era + two key event annotations
#   Note: annotate("richtext") supports <b> and <span> but NOT <a> tags
# ╚══════════════════════════════════════════════════════════════════════════╝

p3 <- ggplot(nafta_section,
             aes(x = year, colour = section, fill = section)) +
  # Shaded USMCA era (post-2020)
  annotate("rect",
           xmin = 2020, xmax = 2025,
           ymin = -Inf, ymax = Inf,
           fill = "#4DA8FF", alpha = 0.06) +
  # USMCA effective date vertical line
  geom_vline(
    xintercept = 2020, linetype = "dashed",
    colour = "#4DA8FF", linewidth = 0.7, alpha = 0.85
  ) +
  # NAFTA phase-out complete marker (~2008)
  geom_vline(
    xintercept = 2008, linetype = "dotted",
    colour = "#7A8FA6", linewidth = 0.5, alpha = 0.7
  ) +
  # IQR uncertainty band
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.20, linewidth = 0) +
  # Mean trajectory
  geom_line(aes(y = mean_rate), linewidth = 1.15) +
  # Small dots for readability at each data point
  geom_point(aes(y = mean_rate),
             size = 1.5, shape = 21, fill = "#141E2B", stroke = 0.9) +
  # USMCA annotation — plain span only, no <a> tags
  annotate("richtext",
           x = 2020.4, y = Inf, hjust = 0, vjust = 1.3,
           label = "<span style='color:#4DA8FF;font-size:7pt'><b>USMCA</b> effective</span>",
           fill = NA, label.colour = NA, size = 2.9) +
  # Phase-out complete annotation
  annotate("richtext",
           x = 2008.3, y = Inf, hjust = 0, vjust = 1.3,
           label = "<span style='color:#7A8FA6;font-size:6.5pt'>Phase-outs complete \u2193</span>",
           fill = NA, label.colour = NA, size = 2.6) +
  scale_colour_manual(values = pal_dark, name = NULL) +
  scale_fill_manual(values   = pal_dark, name = NULL) +
  scale_x_continuous(breaks = seq(2000, 2024, 8)) +
  scale_y_continuous(
    labels = label_number(suffix = "%"),
    limits = c(0, NA)
  ) +
  labs(
    title    = "\u2462 NAFTA \u2192 USMCA TRANSITION",
    subtitle = "Preferential rate mean \u00b1 IQR \u00b7 Rates near 0% = phase-out complete by ~2008",
    x = NULL,
    y = "Avg preferential rate"
  ) +
  theme_dash() +
  theme(legend.position = "none")

# ╔══════════════════════════════════════════════════════════════════════════╗
#   PANEL 4 — BUBBLE: AGREEMENT COVERAGE vs MFN RATE
#   Weighted OLS trend + repelled chapter labels + Ch.24 callout
#   FIX: segment.size (not segment.linewidth) for ggrepel compatibility
# ╚══════════════════════════════════════════════════════════════════════════╝

p4 <- ggplot(bubble_chapter,
             aes(x = n_agreements, y = mean_mfn,
                 size = n_products, colour = section)) +
  # Weighted OLS trend line (weight = number of tariff lines per chapter)
  geom_smooth(
    aes(group = 1, weight = n_products),
    method   = "lm", se = FALSE,
    colour   = "#4DA8FF", linewidth = 0.65,
    linetype = "dotted", alpha = 0.7
  ) +
  # Bubble points
  geom_point(alpha = 0.80) +
  # Chapter labels — segment.size used for ggrepel < 0.9.2 compatibility
  # If your ggrepel >= 0.9.2 you may also use segment.linewidth = 0.4
  geom_text_repel(
    aes(label = chapter_label),
    size          = 2.4,
    max.overlaps  = 20,
    colour        = "#C8D8E8",
    segment.colour = "#2A3F55",
    segment.size  = 0.4,          # ← compatible with all ggrepel versions
    box.padding   = 0.35,
    point.padding = 0.2
  ) +
  # Ch.24 (tobacco) explicit callout — highest MFN in dataset (43.7%)
  # plain span only — gridtext does not support <a> href tags
  annotate("richtext",
           x = 2.4, y = 41,
           label = paste0(
             "<span style='color:#4DA8FF'>",
             "<b>Ch.24 Tobacco</b><br>",
             "43.7% MFN \u00b7 13.1% Bilateral</span>"
           ),
           fill = "#141E2B", label.colour = "#1E2D3D",
           size = 2.7, hjust = 0) +
  scale_size_area(max_size = 12, guide = "none") +
  scale_colour_manual(values = pal_dark, guide = "none") +
  scale_x_continuous(breaks = 0:4, minor_breaks = NULL) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  labs(
    title    = "\u2463 AGREEMENT COVERAGE vs MFN RATE",
    subtitle = "Bubble = HTS chapter \u00b7 Size = tariff lines \u00b7 Dotted = weighted OLS trend",
    x = "Avg preferential agreement groups",
    y = "Mean MFN tariff rate"
  ) +
  theme_dash()

# ╔══════════════════════════════════════════════════════════════════════════╗
#   ASSEMBLE — patchwork 2x2 grid + global title / subtitle / copyright
#   plot.caption uses element_markdown() so keep copyright_str tag-free
# ╚══════════════════════════════════════════════════════════════════════════╝

(p1 + p2 + p3 + p4) +
  plot_layout(ncol = 2, nrow = 2) +
  plot_annotation(
    # Main dashboard title
    title    = "US Agricultural Tariffs \u00b7 #TidyTuesday 2026 Week 17",
    # One-line data provenance
    subtitle = paste0(
      "USITC Tariff Database \u00b7 HTS Chapters 1\u201324 \u00b7 ",
      "461,738 records \u00b7 2,686 tariff lines \u00b7 1997\u20132025 \u00b7 ",
      "Descriptive analysis \u2014 no causal claims"
    ),
    # Copyright + attribution — plain Unicode, NO HTML <a> tags
    caption  = copyright_str,
    theme    = theme(
      # Outer page background matches all panels
      plot.background  = element_rect(fill = "#0F1923", colour = NA),
      # Large bold white title
      plot.title       = element_markdown(
        family = "sans", face = "bold", size = 18,
        colour = "#E8F4FD",
        margin = margin(t = 16, b = 5)
      ),
      # Subtitle — must be element_markdown() to match global theme_tariff class
      # copyright_str contains only plain Unicode — no HTML tags — safe for gridtext
      plot.subtitle    = element_markdown(
        family = "sans", size = 9.5, colour = "#7A8FA6",
        margin = margin(b = 16)
      ),
      # Caption — must be element_markdown() to match global theme_tariff class
      # copyright_str built with glue() using only \u00a9 \u00b7 — no <a> tags
      plot.caption     = element_markdown(
        family = "sans", size = 7.5, colour = "#3E5060",
        hjust  = 0,
        margin = margin(t = 14, b = 12)
      ),
      plot.margin      = margin(8, 18, 8, 18)
    )
  )

# ── Standalone Twitter export (uncomment to run outside Quarto) ───────────
# ggsave(
#   filename = "dashboard_twitter.png",
#   plot     = last_plot(),
#   width    = 18,
#   height   = 11,
#   dpi      = 300,
#   bg       = "#0F1923"
# )
```
```{r}
#| label: tbl-findings
#| tbl-cap: "Seven descriptive findings — #TidyTuesday 2026 Week 17"

tibble(
  `#` = 1:7,
  Finding = c(
    "Live Animals & Animal Products show persistently higher average MFN rates across all years",
    "NAFTA/USMCA agreement codes are associated with lower average rates than MFN across most chapters",
    "Prepared Foodstuffs show the least year-to-year rate variability in the dataset",
    "Rate volatility (SD over time) is concentrated in ~30 tariff lines out of thousands",
    "The 2020 NAFTA → USMCA transition coincides with no discontinuous jump in the rate series",
    "Chapters with higher MFN rates tend to have more preferential agreement groups (descriptive association)",
    "MFN rate distributions are statistically distinguishable across sections (Kruskal-Wallis, Dunn tests)"
  ),
  `Practice area` = c(
    "Data tidying + line/ribbon charts",
    "Pivoting + dumbbell charts",
    "Summarisation + heatmaps",
    "Ranking + lollipop charts",
    "Date handling + annotated time series",
    "Joining datasets + bubble charts",
    "Statistical testing with nonparametric methods"
  )
) |>
  gt() |>
  tab_header(
    title    = md("**Descriptive Findings — #TidyTuesday 2026 W17**"),
    subtitle = md("*Data tidying and plotting practice — no causal claims*")
  ) |>
  cols_label(`#` = "#", `Practice area` = "Technique practised") |>
  cols_width(`#` ~ px(28), Finding ~ px(330), `Practice area` ~ px(240)) |>
  tab_style(style = cell_fill(color = "#EEEDFE"),
            locations = cells_body(rows = seq(1, 7, 2))) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_options(table.font.size = px(13),
              column_labels.border.top.color = "#534AB7",
              column_labels.border.top.width = px(2))
```

---

## How to Share This Work {#sec-share}

<div class="share-box">

**📣 Sharing your #TidyTuesday work**

Following the [TidyTuesday sharing guidelines](https://github.com/rfordatascience/tidytuesday/blob/main/sharing.md):

1. Post your output with **`#TidyTuesday`** on social media
2. Tag **`@R4DSCommunity`** and **`@eyegonk`** (Twitter/X or Mastodon)
3. Share your source code on GitHub
4. Link to the dataset: <https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-04-28/>

**Suggested post text:**
> 📊 My #TidyTuesday Week 17 2026 submission by @eyegonk — exploring US agricultural tariffs from the USITC database. MFN trajectories, NAFTA→USMCA transitions, Lorenz curves, statistical tests across HTS Ch. 1–24. Code: [your link] 🌾 #rstats #dataviz

**Dataset citation:**
> USITC Tariff Database via TidyTuesday 2026, Week 17.
> <https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-04-28/readme.md>

**Rendered with:** [Quarto](https://quarto.org/) · R `r R.version$major`.`r R.version$minor` · ggplot2 · plotly · gt · crosstalk · patchwork

**Author:** Erick K Yegon, PhD · [keyegon@gmail.com](mailto:keyegon@gmail.com) · [@eyegonk](https://twitter.com/eyegonk) · [ORCID 0000-0002-7055-4848](https://orcid.org/0000-0002-7055-4848)

</div>

---data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAABIAAAASCAYAAABWzo5XAAAAbElEQVR4Xs2RQQrAMAgEfZgf7W9LAguybljJpR3wEse5JOL3ZObDb4x1loDhHbBOFU6i2Ddnw2KNiXcdAXygJlwE8OFVBHDgKrLgSInN4WMe9iXiqIVsTMjH7z/GhNTEibOxQswcYIWYOR/zAjBJfiXh3jZ6AAAAAElFTkSuQmCC

## Reproducibility {#sec-repro}

```{r}
#| label: session-info
#| collapse: true
#| code-fold: true
#| code-summary: "Session info"

sessionInfo()
```

---

*Rendered with [Quarto](https://quarto.org) on `r Sys.Date()` by **Erick K Yegon, PhD** · [@eyegonk](https://twitter.com/eyegonk) · [TidyTuesday 2026 Week 17](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-04-28/readme.md) · Share with **#TidyTuesday***
